---
title: "M1 — Analysis template (R)"
subtitle: "Strain gauges and load cell"
description: "In-amp sizing, load-cell calibration, bridge configurations and temperature compensation, in R"
engine: knitr
toc: true
toc-depth: 3
code-fold: false
---

This template runs end to end on the reference data shipped with the
experiment, so you can read the whole analysis before you touch a bench. To
analyse your own measurements, replace the files in `data/` and change the
paths in the first chunk — nothing else.

The Python template [`analysis-py.qmd`](analysis-py.qmd) has the same
sections, the same figure labels and the same output tables, and produces
the same numbers from the same data.

::: {.callout-note collapse="true"}
## Why the creep fit is written out rather than called from a library

The creep model of [the handbook](handbook.qmd#eq-m1-creep) is nonlinear in
its time constant. Both templates fit it with an explicit two-parameter
Gauss–Newton loop, the same discipline
[S3 uses](../S3-aliasing-filtering/analysis-r.qmd) for its sine fit and its
FIR filter: `scipy.optimize.curve_fit` and R's `nls` use different
convergence criteria and can converge to slightly different parameter
values on the same data, which breaks table-for-table agreement between the
two languages.
:::

## Setup

```{r}
#| label: setup
#| message: false
library(ggplot2)
library(patchwork)

source("../../shared/R/labtools.R")

theme_set(theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.25)))

LOAD_MASSES_G <- c(0, 100, 200, 300, 400)
LOADING_RUNS <- sprintf("data/load-loading-%03d.csv", LOAD_MASSES_G)
UNLOADING_RUNS <- sprintf("data/load-unloading-%03d.csv", rev(LOAD_MASSES_G))
CREEP_RUN <- "data/load-creep.csv"
ZERO_RETURN_RUN <- "data/load-zero-return.csv"

BRIDGE_QUARTER_MASSES_G <- c(0, 10, 20, 30)
BRIDGE_QUARTER_RUNS <- sprintf("data/bridge-quarter-%03d.csv", BRIDGE_QUARTER_MASSES_G)
BRIDGE_HALF_RUNS <- c("data/bridge-half-000.csv", "data/bridge-half-030.csv")
BRIDGE_FULL_RUNS <- c("data/bridge-full-000.csv", "data/bridge-full-030.csv")

TEMP_RUNS <- list(
  quarter_cold = "data/temp-quarter-cold.csv",
  quarter_hot = "data/temp-quarter-hot.csv",
  half_cold = "data/temp-half-cold.csv",
  half_hot = "data/temp-half-hot.csv"
)
```

```{r}
#| label: helpers
# Ordinary least squares, y = a + b x, written out so the sensitivity
# coefficients and the residual standard error match the handbook exactly --
# the same construction as S1's calibration fit.
ols_fit <- function(x, y) {
  n <- length(x)
  sxx <- sum((x - mean(x))^2)
  b <- sum((x - mean(x)) * (y - mean(y))) / sxx
  a <- mean(y) - b * mean(x)
  resid <- y - (a + b * x)
  if (n > 2) {
    s_res <- sqrt(sum(resid^2) / (n - 2))
    u_b <- s_res / sqrt(sxx)
    u_a <- s_res * sqrt(1 / n + mean(x)^2 / sxx)
  } else {
    u_b <- NA_real_
    u_a <- NA_real_
  }
  list(a = a, b = b, u_a = u_a, u_b = u_b, resid = resid)
}

# Gauss-Newton fit of y(t) = C (1 - exp(-t/tau)) to y = v/v0 - 1. Two
# parameters only -- v0 is taken directly from the first samples of the
# hold rather than fitted, which keeps the iteration well-conditioned and
# gives Python and R the same starting point.
creep_fit <- function(t_s, v, v0, n_iter = 200) {
  y <- v / v0 - 1
  c_par <- y[length(y)]
  tau <- t_s[length(t_s)] / 3
  for (i in seq_len(n_iter)) {
    e <- exp(-t_s / tau)
    model <- c_par * (1 - e)
    resid <- y - model
    d_c <- 1 - e
    d_tau <- -c_par * e * (t_s / tau^2)
    jac <- cbind(d_c, d_tau)
    step <- qr.solve(jac, resid)
    c_par <- c_par + step[1]
    tau <- tau + step[2]
    if (max(abs(step)) < 1e-12) break
  }
  resid <- y - c_par * (1 - exp(-t_s / tau))
  list(c_inf = c_par, tau_s = tau, resid = resid)
}
```

```{r}
#| label: load
quarter_runs <- lapply(BRIDGE_QUARTER_RUNS, read_run)
primary <- quarter_runs[[length(quarter_runs)]]
board <- primary$meta$board
ch0 <- primary$meta$channels[[1]]

FSR <- board$vref_mV / 1000
LSB <- lsb_volts(primary$meta)
GAUGE_FACTOR <- ch0$gauge_factor
U_GAUGE_FACTOR_REL <- ch0$u_gauge_factor_rel
VEX <- ch0$bridge_excitation_v
U_EXCITATION_REL <- ch0$u_excitation_rel
GAIN_BEAM <- ch0$conditioning$gain
OFFSET_BEAM <- ch0$conditioning$offset
U_GAIN_REL <- ch0$amplifier_u_gain_rel
U_OFFSET_V <- ch0$amplifier_u_offset_v
```

```{r}
#| label: banner
#| echo: false
#| output: asis
cat(synthetic_banner(primary$meta))
```

## Data quality

```{r}
#| label: quality
#| echo: false
#| output: asis
all_paths <- c(LOADING_RUNS, UNLOADING_RUNS, CREEP_RUN, ZERO_RETURN_RUN,
               BRIDGE_QUARTER_RUNS, BRIDGE_HALF_RUNS, BRIDGE_FULL_RUNS,
               unlist(TEMP_RUNS))
all_runs <- lapply(all_paths, read_run)
names(all_runs) <- all_paths
n_bad <- sum(vapply(all_runs, function(r) length(validate_meta(r$meta)) > 0, logical(1)))
lossy <- Filter(function(name) {
  stats <- all_runs[[name]]$meta$stream_stats
  counters <- stats[setdiff(names(stats), "records")]
  any(vapply(counters, function(v) is.numeric(v) && v > 0, logical(1)))
}, names(all_runs))
cat(sprintf('
%d runs read.
%s
%s
', length(all_runs),
   if (n_bad == 0) "All sidecars are complete."
   else sprintf("%d sidecar(s) have missing fields.", n_bad),
   if (length(lossy) == 0) "No run reports a lost record."
   else paste0("Runs with losses: ", paste(lossy, collapse = ", "))))
```

## Why the amplifier: the arithmetic

At a representative strain $\varepsilon = 10^{-4}$
([the handbook](handbook.qmd#eq-m1-quarter-numeric)), the bare quarter-bridge
signal, the LSB it would be read against with no amplifier, and the actual
amplified signal this bench uses, computed from the sidecar's own recorded
gauge factor, excitation and gain — not re-typed constants.

```{r}
#| label: inamp
EPS_EXAMPLE <- 1e-4
raw_v <- GAUGE_FACTOR * EPS_EXAMPLE * VEX / 4
raw_lsb <- raw_v / LSB
amplified_v <- GAIN_BEAM * raw_v
amplified_frac_fs <- amplified_v / FSR
```

```{r}
#| label: out-inamp
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| Representative strain $\\varepsilon$ | %.0e |
| Bare bridge signal, $k\\varepsilon V_\\mathrm{ex}/4$ | %.4f mV |
| One LSB on this board | %.2f µV |
| Bare signal, in LSB | %.2f |
| Amplifier gain actually used | %.1f |
| Amplified signal at the ADC | %.3f V |
| Amplified signal, fraction of full scale | %.1f %% |

: The case for the amplifier, in numbers {#tbl-m1-inamp}
', EPS_EXAMPLE, raw_v * 1e3, LSB * 1e6, raw_lsb, GAIN_BEAM, amplified_v,
   amplified_frac_fs * 100))
```

```{r}
#| label: out-inamp-text
#| echo: false
#| output: asis
cat(sprintf('
Unamplified, this strain would produce **%.1f LSB** — the same
order of magnitude as [S1\'s own noise floor](../S1-daq-static/handbook.qmd#sec-s1-enob),
which on this board is of order 0.5 mV RMS, itself several LSB wide. A
bridge signal that size is not a small measurement; it is not a measurement.
Amplified by %.0f, the same strain reaches
**%.0f %%** of the ADC\'s full-scale range, which is
what [the required-gain design](handbook.qmd#eq-m1-required-gain) is for.
', raw_lsb, GAIN_BEAM, amplified_frac_fs * 100))
```

## Load cell: linearity and hysteresis

The load cell's static calibration: a straight-line fit to the loading
sweep, and the same sweep read backwards.

```{r}
#| label: linearity
load_frame <- function(paths, direction) {
  rows <- lapply(paths, function(p) {
    run <- read_run(p)
    ch <- run$meta$channels[[1]]
    data.frame(run = run$meta$run, direction = direction,
               F_N = ch$reference_value, u_F_N = ch$reference_u,
               mean_V = mean(run$data$volts), sd_V = sd(run$data$volts),
               n = nrow(run$data), capacity_N = ch$load_cell_capacity_n,
               sensitivity_mv_per_v = ch$load_cell_sensitivity_mv_per_v,
               excitation_v = ch$bridge_excitation_v, gain = ch$conditioning$gain)
  })
  do.call(rbind, rows)
}

loading <- load_frame(LOADING_RUNS, "loading")
unloading <- load_frame(UNLOADING_RUNS, "unloading")

row0 <- loading[1, ]
V_FS <- row0$gain * row0$sensitivity_mv_per_v * 1e-3 * row0$excitation_v

fit <- ols_fit(loading$F_N, loading$mean_V)
loading$residual_V <- fit$resid
nonlinearity_pct <- max(abs(loading$residual_V)) / V_FS * 100
```

```{r}
#| label: out-linearity
#| echo: false
#| output: asis
rows <- sprintf("| %s | %.4f | %.5f | %.3f | %d | %+.3f |",
                loading$run, loading$F_N, loading$mean_V, loading$sd_V * 1e3,
                loading$n, loading$residual_V * 1e3)
cat(sprintf('
| Run | Force [N] | Mean [V] | $s$ [mV] | $n$ | Residual [mV] |
|---|---|---|---|---|---|
%s

: Loading sweep and the linear fit\'s residuals {#tbl-m1-linearity}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: out-linearity-text
#| echo: false
#| output: asis
cat(sprintf('
Fitted $a = %.2f$ mV, $b = %.3f$ mV/N, with
$u(a) = %.3f$ mV and $u(b) = %.4f$ mV/N.
Full-scale output $V_\\mathrm{FS} = %.3f$ V (at the load cell\'s rated
capacity, from the sidecar). Non-linearity, from [the linearity definition in the handbook](handbook.qmd#eq-m1-linearity):
**%.2f %%FS**.
', fit$a * 1e3, fit$b * 1e3, fit$u_a * 1e3, fit$u_b * 1e3, V_FS,
   nonlinearity_pct))
```

```{r}
#| label: fig-m1-linearity
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.0
#| fig-cap: "Loading sweep: indication against force (top) and fit residuals (bottom)."
p_fit <- ggplot(loading, aes(F_N, mean_V)) +
  geom_abline(intercept = fit$a, slope = fit$b, linewidth = 0.5) +
  geom_point(size = 2.4) +
  labs(x = NULL, y = "Indicated voltage [V]")

p_res <- ggplot(loading, aes(F_N, residual_V * 1e3)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.4) +
  geom_point(size = 2.4) +
  labs(x = "Force [N]", y = "Residual [mV]")

p_fit / p_res + plot_layout(heights = c(2, 1))
```

```{r}
#| label: hysteresis
unloading_sorted <- unloading[order(unloading$F_N), ]
hyst <- data.frame(F_N = loading$F_N, loading_V = loading$mean_V,
                   unloading_V = unloading_sorted$mean_V)
hyst$diff_mV <- (hyst$unloading_V - hyst$loading_V) * 1e3
hysteresis_pct <- max(abs(hyst$diff_mV)) / 1e3 / V_FS * 100
```

```{r}
#| label: out-hysteresis
#| echo: false
#| output: asis
rows <- sprintf("| %.4f | %.5f | %.5f | %+.3f |",
                hyst$F_N, hyst$loading_V, hyst$unloading_V, hyst$diff_mV)
cat(sprintf('
| Force [N] | Loading [V] | Unloading [V] | Difference [mV] |
|---|---|---|---|
%s

: Loading against unloading at the same nominal force {#tbl-m1-hysteresis}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: out-hysteresis-text
#| echo: false
#| output: asis
cat(sprintf('
Maximum hysteresis, from [the hysteresis definition in the handbook](handbook.qmd#eq-m1-hysteresis): **%.2f %%FS**,
largest near mid-range and closing towards both ends, as the bell-shaped
path dependence predicts.
', hysteresis_pct))
```

```{r}
#| label: fig-m1-hysteresis
#| echo: false
#| fig-width: 6.5
#| fig-height: 4.0
#| fig-cap: "Loading and unloading curves overlaid. The gap between them at each force is the hysteresis of @tbl-m1-hysteresis."
trace <- rbind(
  data.frame(F_N = hyst$F_N, V = hyst$loading_V, case = "loading"),
  data.frame(F_N = hyst$F_N, V = hyst$unloading_V, case = "unloading"))
ggplot(trace, aes(F_N, V, colour = case)) +
  geom_line(linewidth = 0.4) +
  geom_point(size = 2.2) +
  labs(x = "Force [N]", y = "Indicated voltage [V]", colour = NULL)
```

## Load cell: creep and zero drift

```{r}
#| label: creep
creep_run <- read_run(CREEP_RUN)
t_s <- (creep_run$data$t_us - creep_run$data$t_us[1]) / 1e6
v <- creep_run$data$volts
v0 <- mean(v[1:10])
creep <- creep_fit(t_s, v, v0)
```

```{r}
#| label: out-creep
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| $V(0)$ | %.5f V |
| Fitted $C_\\infty$ | %.2f %% |
| Fitted $\\tau$ | %.1f s |
| Residual RMS | %.3f mV (relative units) |
| Hold duration | %.0f s |

: Creep fit of [the creep model in the handbook](handbook.qmd#eq-m1-creep) on the held load {#tbl-m1-creep}
', v0, creep$c_inf * 100, creep$tau_s, sqrt(mean(creep$resid^2)) * 1e3,
   t_s[length(t_s)]))
```

```{r}
#| label: fig-m1-creep
#| echo: false
#| fig-width: 6.5
#| fig-height: 3.6
#| fig-cap: "The held-load reading over time, with the fitted creep curve of [the creep model in the handbook](handbook.qmd#eq-m1-creep)."
fit_curve <- data.frame(t_s = t_s,
                        v = v0 * (1 + creep$c_inf * (1 - exp(-t_s / creep$tau_s))))
ggplot() +
  geom_line(data = data.frame(t_s = t_s, v = v), aes(t_s, v),
            linewidth = 0.25, alpha = 0.6) +
  geom_line(data = fit_curve, aes(t_s, v), linewidth = 0.6, colour = "#c0392b") +
  labs(x = "Time [s]", y = "Indicated voltage [V]")
```

```{r}
#| label: zero_drift
first_zero <- mean(read_run(LOADING_RUNS[1])$data$volts)
last_zero <- mean(read_run(ZERO_RETURN_RUN)$data$volts)
zero_drift_mV <- (last_zero - first_zero) * 1e3
zero_drift_pct <- abs(zero_drift_mV) / 1e3 / V_FS * 100
```

```{r}
#| label: out-zero
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| First zero (start of session) | %.5f V |
| Last zero (after the full sequence) | %.5f V |
| Zero drift | %+.3f mV (%.3f %%FS) |

: Zero drift across the session {#tbl-m1-zero}
', first_zero, last_zero, zero_drift_mV, zero_drift_pct))
```

## Bridge configurations on the beam

The same small reference masses, read through quarter, half and full
bridge. The quarter-bridge points also give the load-to-strain calibration
[M2](../M2-hall-displacement/index.qmd) reads back.

```{r}
#| label: bridges
bridge_points <- function(paths) {
  rows <- lapply(paths, function(p) {
    run <- read_run(p)
    ch <- run$meta$channels[[1]]
    data.frame(F_N = ch$reference_value, mean_V = mean(run$data$volts),
               sd_V = sd(run$data$volts), n = nrow(run$data))
  })
  df <- do.call(rbind, rows)
  df[order(df$F_N), ]
}

quarter <- bridge_points(BRIDGE_QUARTER_RUNS)
half <- bridge_points(BRIDGE_HALF_RUNS)
full <- bridge_points(BRIDGE_FULL_RUNS)

fit_quarter <- ols_fit(quarter$F_N, quarter$mean_V)
slope_quarter <- fit_quarter$b
u_slope_quarter <- fit_quarter$u_b
slope_half <- (half$mean_V[2] - half$mean_V[1]) / (half$F_N[2] - half$F_N[1])
slope_full <- (full$mean_V[2] - full$mean_V[1]) / (full$F_N[2] - full$F_N[1])

ratio_half <- slope_half / slope_quarter
ratio_full <- slope_full / slope_quarter

# Strain sensitivity at the gauge, recovered from the quarter-bridge slope:
# V_adc = GAIN_BEAM * (k eps Vex / 4) + offset, eps linear in F, so
# d(eps)/dF = 4 * d(V_adc)/dF / (GAIN_BEAM k Vex).
d_eps_d_f <- 4 * slope_quarter / (GAIN_BEAM * GAUGE_FACTOR * VEX)
u_d_eps_d_f <- 4 * u_slope_quarter / (GAIN_BEAM * GAUGE_FACTOR * VEX)
```

```{r}
#| label: out-bridges
#| echo: false
#| output: asis
cat(sprintf('
| Configuration | $dV/dF$ [mV/N] | Ratio to quarter | Nominal ratio |
|---|---|---|---|
| Quarter | %.3f | 1.00 | 1 |
| Half | %.3f | %.3f | 2 |
| Full | %.3f | %.3f | 4 |

: Bridge sensitivity and the 1 : 2 : 4 ratio of [the bridge relations in the handbook](handbook.qmd#eq-m1-bridge-configs) {#tbl-m1-bridges}
', slope_quarter * 1e3, slope_half * 1e3, ratio_half, slope_full * 1e3, ratio_full))
```

```{r}
#| label: out-bridges-text
#| echo: false
#| output: asis
cat(sprintf('
Half and full bridge reproduce the %.2f× and %.2f×
sensitivity [the bridge relations in the handbook](handbook.qmd#eq-m1-bridge-configs) predicts. The half- and full-bridge
sensitivities come from only two points each — enough to confirm the ratio,
not enough to give a fit its own residual, the same limitation
[S1 notes](../S1-daq-static/handbook.qmd#sec-s1-instructor-failures) for a
two-point calibration.

From the quarter-bridge slope, the strain sensitivity to force at the gauge
is $d\\varepsilon/dF = %.3f \\times 10^{-6}$ /N, with
$u(d\\varepsilon/dF) = %.4f \\times 10^{-6}$ /N — the
number [M2](../M2-hall-displacement/analysis-r.qmd) reads back for its
stiffness cross-check.
', ratio_half, ratio_full, d_eps_d_f * 1e6, u_d_eps_d_f * 1e6))
```

```{r}
#| label: fig-m1-bridges
#| echo: false
#| fig-width: 6.5
#| fig-height: 4.0
#| fig-cap: "Bridge output against applied force for the three configurations. The slopes are in the ratio 1 : 2 : 4."
bridges_long <- rbind(
  data.frame(F_N = quarter$F_N, V = quarter$mean_V, config = "quarter"),
  data.frame(F_N = half$F_N, V = half$mean_V, config = "half"),
  data.frame(F_N = full$F_N, V = full$mean_V, config = "full"))
ggplot(bridges_long, aes(F_N, V, colour = config)) +
  geom_line(linewidth = 0.4) +
  geom_point(size = 2.2) +
  labs(x = "Force [N]", y = "Indicated voltage [V]", colour = NULL)
```

## Temperature compensation

Beam unloaded, quarter bridge against a dummy-compensated half bridge, at
two bench temperatures.

```{r}
#| label: temperature
temp <- lapply(TEMP_RUNS, read_run)
delta_t <- temp$quarter_hot$meta$channels[[1]]$applied_delta_t_c

apparent_alpha <- function(cold_key, hot_key, n_arm) {
  v_cold <- mean(temp[[cold_key]]$data$volts)
  v_hot <- mean(temp[[hot_key]]$data$volts)
  d_v_bridge <- (v_hot - v_cold) / GAIN_BEAM   # offset cancels in the difference
  (4 * d_v_bridge / VEX) / (n_arm * GAUGE_FACTOR) / delta_t
}

alpha_quarter <- apparent_alpha("quarter_cold", "quarter_hot", 1)
alpha_half <- apparent_alpha("half_cold", "half_hot", 2)
compensation_ratio <- alpha_quarter / alpha_half
```

```{r}
#| label: out-temp
#| echo: false
#| output: asis
cat(sprintf('
| Configuration | Apparent strain / K |
|---|---|
| Quarter bridge (bare) | $%.3f \\times 10^{-6}$ /K |
| Half bridge (dummy-compensated) | $%.4f \\times 10^{-6}$ /K |

: Apparent strain per kelvin, from [the apparent-strain relation in the handbook](handbook.qmd#eq-m1-temp-apparent), $\\Delta T = %g$ C {#tbl-m1-temp}
', alpha_quarter * 1e6, alpha_half * 1e6, delta_t))
```

```{r}
#| label: out-temp-text
#| echo: false
#| output: asis
cat(sprintf('
The dummy-compensated half bridge\'s apparent strain per kelvin is
**%.1f times smaller** than the bare quarter bridge\'s —
compensation, not cancellation: the residual is real and belongs in any
budget that uses the half-bridge configuration in a varying-temperature
environment.
', compensation_ratio))
```

```{r}
#| label: fig-m1-temp
#| echo: false
#| fig-width: 5.0
#| fig-height: 3.6
#| fig-cap: "Apparent strain per kelvin, quarter bridge against dummy-compensated half bridge."
bar_data <- data.frame(config = factor(c("quarter", "half"), levels = c("quarter", "half")),
                       alpha = c(alpha_quarter, alpha_half) * 1e6)
ggplot(bar_data, aes(config, alpha)) +
  geom_col(fill = "#2a7fa8") +
  labs(x = NULL, y = "Apparent strain / K  [x 1e-6]")
```

## Uncertainty budget

The measurand is the beam-mounted quarter bridge's strain,
$\hat\varepsilon$ from [the handbook](handbook.qmd#eq-m1-strain-recover), at
the largest quarter-bridge reference point, with the [sensitivity
coefficients](handbook.qmd#eq-m1-sensitivity) given there.

```{r}
#| label: budget
row <- quarter[nrow(quarter), ]
v_adc_mean <- row$mean_V
sigma_v_adc <- row$sd_V
n_pts <- row$n
delta_v <- (v_adc_mean - OFFSET_BEAM) / GAIN_BEAM
eps_hat <- 4 * delta_v / (GAUGE_FACTOR * VEX)

c_dv <- 4 / (GAUGE_FACTOR * VEX)
c_gain <- -eps_hat / GAIN_BEAM
c_offset <- -4 / (GAUGE_FACTOR * VEX * GAIN_BEAM)
c_k <- -eps_hat / GAUGE_FACTOR
c_vex <- -eps_hat / VEX

enob_here <- enob(FSR, sigma_v_adc)
q_eff_adc <- FSR / 2^enob_here
q_eff_bridge <- q_eff_adc / GAIN_BEAM     # referred to Delta V, same units as delta_v

components <- list(
  list(quantity = "Repeatability of Delta V", value = delta_v, unit = "strain",
       distribution = "type-a", std = (sigma_v_adc / GAIN_BEAM) / sqrt(n_pts),
       sensitivity = c_dv),
  list(quantity = "Effective quantization", value = 0, unit = "strain",
       distribution = "rectangular", half_width = q_eff_bridge / 2, sensitivity = c_dv),
  list(quantity = "Amplifier gain G", value = GAIN_BEAM, unit = "strain",
       distribution = "rectangular", half_width = U_GAIN_REL * GAIN_BEAM,
       sensitivity = c_gain),
  list(quantity = "Amplifier offset", value = OFFSET_BEAM, unit = "strain",
       distribution = "rectangular", half_width = U_OFFSET_V, sensitivity = c_offset),
  list(quantity = "Gauge factor k", value = GAUGE_FACTOR, unit = "strain",
       distribution = "rectangular", half_width = U_GAUGE_FACTOR_REL * GAUGE_FACTOR,
       sensitivity = c_k),
  list(quantity = "Excitation voltage Vex", value = VEX, unit = "strain",
       distribution = "rectangular", half_width = U_EXCITATION_REL * VEX,
       sensitivity = c_vex)
)
budget <- budget_table(components, k = 2)
```

```{r}
#| label: out-budget
#| echo: false
#| output: asis
cat(budget_to_markdown(
  budget,
  caption = sprintf("Uncertainty budget for the recovered strain at $F = %.3f$ N", row$F_N),
  label = "tbl-m1-budget"))
```

```{r}
#| label: out-budget-text
#| echo: false
#| output: asis
dominant <- budget$table[which.max(budget$table$index_pct), ]
cat(sprintf('
Recovered strain $\\hat\\varepsilon = %.3f \\times 10^{-6}$,
with expanded uncertainty $U = %.3f \\times 10^{-6}$ at
$k = 2$ — a relative uncertainty of %.1f %%. The
dominant component is **%s** at
%.0f %% of the combined variance.
', eps_hat * 1e6, budget$U * 1e6, budget$U / eps_hat * 100, dominant$quantity,
   dominant$index_pct))
```

::: {.callout-note}
## What this budget assumes

Inputs are treated as uncorrelated, and the model is linear at the operating
point — both standard simplifications, stated here as
[the shared uncertainty-budget guidance](../../data-format/index.qmd#sec-budget)
asks. The load cell's own hysteresis and creep are not included, because the
measurand budgeted here is the beam-mounted gauge's strain, not the load
cell's force; see [the handbook](handbook.qmd#sec-m1-budget) for why.
:::

## What to report

1. @tbl-m1-inamp: the in-amp arithmetic, with your own sidecar's gauge
   factor and gain.
2. @tbl-m1-linearity and @fig-m1-linearity: the loading fit and its
   residuals, with non-linearity as %FS.
3. @tbl-m1-hysteresis and @fig-m1-hysteresis: the loading/unloading loop,
   with maximum hysteresis as %FS.
4. @tbl-m1-creep and @fig-m1-creep: the fitted creep curve.
5. @tbl-m1-zero: zero drift across the session.
6. @tbl-m1-bridges and @fig-m1-bridges: the 1 : 2 : 4 sensitivity ratio.
7. @tbl-m1-temp and @fig-m1-temp: apparent strain per kelvin, both
   configurations, and the compensation ratio between them.
8. @tbl-m1-budget: the budget, the dominant component, and what you would
   change first to reduce it.
