---
title: "T1 — Analysis template (R)"
subtitle: "Static calibration of temperature sensors"
description: "NTC and Pt1000 calibration, sensitivity, resolution and self-heating, in R"
engine: knitr
toc: true
toc-depth: 3
code-fold: false
---

This template runs end to end on the reference data shipped with the
experiment, so you can read the whole analysis before you touch a board. To
analyse your own measurements, replace the files in `data/` and change the
paths in the first chunk — nothing else.

The Python template [`analysis-py.qmd`](analysis-py.qmd) has the same
sections, the same figure labels and the same output tables, and produces the
same numbers from the same data.

::: {.callout-note collapse="true"}
## Why the sensitivity derivatives are computed by finite difference

Both sensor calibrations are genuinely nonlinear, and this template needs
derivatives of the fitted curves at an operating point — for
[the sensitivity comparison](handbook.qmd#sec-t1-sensitivity) and for
[the budget's sensitivity coefficient](handbook.qmd#eq-t1-budget-sens). A hand-derived
closed form, transcribed once into Python and once into R, is exactly the
kind of place the two could silently disagree. Both templates instead take
the same central difference of the same fitted function, which by
construction computes the same number in both languages.
:::

## Setup

```{r}
#| label: setup
#| message: false
library(ggplot2)
library(purrr)
library(patchwork)

source("../../shared/R/labtools.R")

CAL_RUNS <- c(ice = "data/cal-ice.csv", t20 = "data/cal-t20.csv",
             t35 = "data/cal-t35.csv", t50 = "data/cal-t50.csv",
             t65 = "data/cal-t65.csv", t80 = "data/cal-t80.csv")
SELFHEAT_RUN <- "data/selfheat-high.csv"
S1_NOISE_RUN <- "../S1-daq-static/data/noise-linux.csv"

PIN_NTC <- 15
PIN_PT <- 16
OPERATING_POINT <- "t50"     # the bath point reused for the self-heating run

# IEC 60751 Callendar-Van Dusen coefficients, t >= 0 degC branch: the
# published standard, fixed rather than fitted (see _theory.qmd).
A_IEC <- 3.9083e-3
B_IEC <- -5.775e-7
```

```{r}
#| label: helpers
# (f(x+eps) - f(x-eps)) / (2 eps) -- the one derivative rule both templates
# share, so a fitted curve's local slope is the same number in both
# languages regardless of how messy its closed form would be.
central_diff <- function(f, x, eps) (f(x + eps) - f(x - eps)) / (2 * eps)

# Inverts the divider of @eq-t1-divider for the NTC's resistance.
r_ntc_from_v <- function(v, r_fixed, vcc) vcc * r_fixed / v - r_fixed

# Steinhart-Hart, @eq-t1-sh, evaluated from a fitted (A, B, C).
t_ntc_from_r <- function(r, coef) {
  ln_r <- log(r)
  inv_t <- coef[1] + coef[2] * ln_r + coef[3] * ln_r^3
  1 / inv_t - 273.15
}

# Inverts the bridge of @eq-t1-bridge for the Pt1000's resistance.
r_pt_from_v <- function(v, r_bridge, v_exc, gain, offset) {
  deflection <- (v - offset) / (gain * v_exc)
  r_bridge * (0.5 + deflection) / (0.5 - deflection)
}

# Callendar-Van Dusen, @eq-t1-cvd, forward direction.
r_pt_from_t <- function(t, r0, a, b = B_IEC) r0 * (1 + a * t + b * t^2)

# Callendar-Van Dusen, positive root of the quadratic in t.
t_pt_from_r <- function(r, r0, a, b = B_IEC) {
  c_coef <- 1 - r / r0
  (-a + sqrt(a^2 - 4 * b * c_coef)) / (2 * b)
}

# The channel entry for one pin, from a run's sidecar.
channel_for_pin <- function(run, pin) {
  Filter(function(c) c$pin == pin, run$meta$channels)[[1]]
}
```

```{r}
#| label: load
cal <- map(CAL_RUNS, read_run)
selfheat <- read_run(SELFHEAT_RUN)
primary <- cal[[OPERATING_POINT]]
board <- primary$meta$board
LSB <- lsb_volts(primary$meta)
FSR <- board$vref_mV / 1000
```

```{r}
#| label: banner
#| echo: false
#| output: asis
cat(synthetic_banner(primary$meta))
```

## Data quality

```{r}
#| label: quality
#| echo: false
#| output: asis
all_runs <- c(cal, list(`selfheat-high` = selfheat))
for (name in names(all_runs)) {
  meta <- all_runs[[name]]$meta
  problems <- validate_meta(meta)
  stats <- meta$stream_stats
  counters <- stats[setdiff(names(stats), "records")]
  lossy <- counters[map_lgl(counters, ~ is.numeric(.x) && .x > 0)]
  cat("\n**", name, "** — ", stats$records, " records; ",
      if (length(problems) == 0) "metadata complete"
      else paste0("metadata problems: ", paste(problems, collapse = "; ")),
      "; ",
      if (length(lossy) == 0) "no losses reported"
      else paste0("losses: ",
                  paste(names(lossy), unlist(lossy), sep = " = ", collapse = ", ")),
      "\n", sep = "")
}
```

## Correcting for the ADC's own offset and gain

S1 measured this platform's own ADC offset and gain error directly, from a
three-point calibration against a DMM
([the calibration model](../S1-daq-static/handbook.qmd#eq-s1-fit)). That
correction is not optional here: the front ends below produce voltages that
differ by close to an order of magnitude between the coldest calibration
point and the self-heating run's high excitation, so a *fixed* voltage error
is a very different *fraction* of the reading at each one, and inverting an
uncorrected voltage through either sensor's strongly nonlinear resistance
relation turns that fraction into a resistance error that does not cancel
between conditions. Every voltage used below is corrected with S1's own
fitted offset and gain first.

```{r}
#| label: s1-correction
s1_paths <- c("../S1-daq-static/data/cal-0v.csv",
             "../S1-daq-static/data/cal-1v25.csv",
             "../S1-daq-static/data/cal-2v50.csv")
s1_points <- map(s1_paths, function(path) {
  run <- read_run(path)
  pin <- run$meta$acquisition$pins[[1]]
  ch <- channel_for_pin(run, pin)
  volts <- run$data$volts[run$data$pin == pin]
  data.frame(x = ch$reference_value, y = mean(volts))
}) |> list_rbind()
sxx_s1 <- sum((s1_points$x - mean(s1_points$x))^2)
B_S1 <- sum((s1_points$x - mean(s1_points$x)) * (s1_points$y - mean(s1_points$y))) / sxx_s1
A_S1 <- mean(s1_points$y) - B_S1 * mean(s1_points$x)

# Removes S1's fitted ADC offset and gain error from a raw voltage.
s1_correct <- function(v) (v - A_S1) / B_S1
```

## NTC calibration: Steinhart–Hart

Each bath point gives a mean divider voltage on pin 15, corrected as above
and inverted to a resistance with the design values from that run's own
sidecar, then fit to [the Steinhart–Hart equation](handbook.qmd#eq-t1-sh-linear).

```{r}
#| label: ntc-calibration
ntc <- map(cal, function(run) {
  ch <- channel_for_pin(run, PIN_NTC)
  v <- s1_correct(mean(run$data$volts[run$data$pin == PIN_NTC]))
  r <- r_ntc_from_v(v, ch$divider_r_fixed_ohm, ch$excitation_v)
  data.frame(run = run$meta$run, bath_C = ch$reference_value,
             u_bath_C = ch$reference_u, R_ohm = r, mean_V = v,
             n = sum(run$data$pin == PIN_NTC))
}) |> list_rbind()
ntc <- ntc[order(ntc$bath_C), ]
rownames(ntc) <- NULL

# Ordinary least squares of 1/T on (1, ln R, (ln R)^3): @eq-t1-sh-linear.
ln_r <- log(ntc$R_ohm)
y <- 1 / (ntc$bath_C + 273.15)
X <- cbind(1, ln_r, ln_r^3)
coef_sh <- solve(crossprod(X), crossprod(X, y))[, 1]

ntc$T_fit_C <- t_ntc_from_r(ntc$R_ohm, coef_sh)
ntc$residual_K <- ntc$T_fit_C - ntc$bath_C
```

```{r}
#| label: out-ntc-cal
#| echo: false
#| output: asis
rows <- sprintf("| %s | %.2f | %.3f | %.2f | %.3f | %+.1f |",
                ntc$run, ntc$bath_C, ntc$u_bath_C, ntc$R_ohm, ntc$T_fit_C,
                ntc$residual_K * 1e3)
cat(sprintf('
| Run | Bath (degC) | $u$(bath) (K) | $R$ (ohm) | Fitted $T$ (degC) | Residual (mK) |
|---|---|---|---|---|---|
%s

: NTC calibration points and Steinhart-Hart fit residuals {#tbl-t1-ntc-cal}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: out-ntc-cal-text
#| echo: false
#| output: asis
cat(sprintf('
Fitted Steinhart-Hart coefficients $A = %.6e$,
$B = %.6e$, $C = %.6e$. Largest residual
$%.1f$ mK — the bound that enters the
budget as fit residual, because six points cannot correct a curve shape they
did not sample.
', coef_sh[1], coef_sh[2], coef_sh[3], max(abs(ntc$residual_K)) * 1e3))
```

```{r}
#| label: fig-t1-ntc-cal
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.2
#| fig-cap: "NTC calibration: bath temperature against resistance (top), with the fitted Steinhart-Hart curve, and fit residuals in millikelvin (bottom)."
r_grid <- seq(min(ntc$R_ohm) * 0.95, max(ntc$R_ohm) * 1.05, length.out = 200)
fit_curve <- data.frame(R_ohm = r_grid, T_C = t_ntc_from_r(r_grid, coef_sh))

p_fit <- ggplot() +
  geom_line(data = fit_curve, aes(R_ohm, T_C), linewidth = 0.5) +
  geom_point(data = ntc, aes(R_ohm, bath_C), size = 2.4) +
  labs(x = "Resistance (ohm)", y = "Temperature (degC)")

p_res <- ggplot(ntc, aes(bath_C, residual_K * 1e3)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.4) +
  geom_point(size = 2.4) +
  labs(x = "Bath temperature (degC)", y = "Residual (mK)")

p_fit / p_res + plot_layout(heights = c(2, 1))
```

## Pt1000 calibration: Callendar–Van Dusen

Same idea, pin 16, but the fit recovers $R_0$ and $A$ with
[$B$ fixed at the IEC 60751 value](handbook.qmd#eq-t1-cvd-linear).

```{r}
#| label: pt-calibration
pt <- map(cal, function(run) {
  ch <- channel_for_pin(run, PIN_PT)
  v <- s1_correct(mean(run$data$volts[run$data$pin == PIN_PT]))
  r <- r_pt_from_v(v, ch$bridge_r_ohm, ch$excitation_v,
                   ch$conditioning$gain, ch$conditioning$offset)
  data.frame(run = run$meta$run, bath_C = ch$reference_value,
             u_bath_C = ch$reference_u, R_ohm = r, mean_V = v,
             n = sum(run$data$pin == PIN_PT))
}) |> list_rbind()
pt <- pt[order(pt$bath_C), ]
rownames(pt) <- NULL

# Linear least squares of R on (1 + B t^2, t): @eq-t1-cvd-linear.
t_bath <- pt$bath_C
basis1 <- 1 + B_IEC * t_bath^2
basis2 <- t_bath
Xp <- cbind(basis1, basis2)
coef_p <- solve(crossprod(Xp), crossprod(Xp, pt$R_ohm))[, 1]
R0_FIT <- coef_p[1]
A_FIT <- coef_p[2] / coef_p[1]

pt$T_fit_C <- t_pt_from_r(pt$R_ohm, R0_FIT, A_FIT)
pt$residual_K <- pt$T_fit_C - pt$bath_C
```

```{r}
#| label: out-pt-cal
#| echo: false
#| output: asis
rows <- sprintf("| %s | %.2f | %.3f | %.3f | %.3f | %+.1f |",
                pt$run, pt$bath_C, pt$u_bath_C, pt$R_ohm, pt$T_fit_C,
                pt$residual_K * 1e3)
cat(sprintf('
| Run | Bath (degC) | $u$(bath) (K) | $R$ (ohm) | Fitted $T$ (degC) | Residual (mK) |
|---|---|---|---|---|---|
%s

: Pt1000 calibration points and Callendar-Van Dusen fit residuals {#tbl-t1-pt-cal}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: out-pt-cal-text
#| echo: false
#| output: asis
cat(sprintf('
Fitted $R_0 = %.3f\\ \\Omega$, fitted $A = %.6e\\ ^\\circ\\mathrm{C}^{-1}$
against the fixed IEC 60751 value $A_\\mathrm{IEC} = %.6e$ — a
difference of %+.2f %%. Largest residual
$%.1f$ mK.
', R0_FIT, A_FIT, A_IEC, (A_FIT / A_IEC - 1) * 100, max(abs(pt$residual_K)) * 1e3))
```

```{r}
#| label: fig-t1-pt-cal
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.2
#| fig-cap: "Pt1000 calibration: resistance against bath temperature (top), with the fitted Callendar-Van Dusen curve, and fit residuals in millikelvin (bottom)."
t_grid <- seq(min(pt$bath_C), max(pt$bath_C), length.out = 200)
fit_curve_pt <- data.frame(bath_C = t_grid, R_ohm = r_pt_from_t(t_grid, R0_FIT, A_FIT))

p_fit_pt <- ggplot() +
  geom_line(data = fit_curve_pt, aes(bath_C, R_ohm), linewidth = 0.5) +
  geom_point(data = pt, aes(bath_C, R_ohm), size = 2.4) +
  labs(x = "Bath temperature (degC)", y = "Resistance (ohm)")

p_res_pt <- ggplot(pt, aes(bath_C, residual_K * 1e3)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.4) +
  geom_point(size = 2.4) +
  labs(x = "Bath temperature (degC)", y = "Residual (mK)")

p_fit_pt / p_res_pt + plot_layout(heights = c(2, 1))
```

## Sensitivity and resolution

[S1](../S1-daq-static/index.qmd) measured this platform's own ADC noise
floor with no sensor attached; the shorted channel of its long run is the
converter's own floor, independent of either front end here. Dividing it by
each front end's sensitivity at the operating point turns a voltage noise
into [a temperature resolution](handbook.qmd#eq-t1-resolution).

```{r}
#| label: sensitivity
op <- cal[[OPERATING_POINT]]
ch_ntc_op <- channel_for_pin(op, PIN_NTC)
ch_pt_op <- channel_for_pin(op, PIN_PT)
T_OP <- ch_ntc_op$reference_value

# The already-corrected resistances from the calibration tables above, not
# recomputed, so the operating point used here is exactly the one that went
# into the fit.
r_ntc_op <- ntc$R_ohm[ntc$run == op$meta$run]
r_pt_op <- pt$R_ohm[pt$run == op$meta$run]

# dR/dT for the NTC: the reciprocal of the fitted model's own dT/dR, taken by
# central difference so no cubic-in-ln(R) has to be inverted by hand.
dT_dR_ntc <- central_diff(function(r) t_ntc_from_r(r, coef_sh), r_ntc_op, 1e-3)
dR_dT_ntc <- 1 / dT_dR_ntc
# dV/dR (divider) times dR/dT: @eq-t1-divider-sens, Vcc*R*Rt/(R+Rt)^2*alpha
# with alpha = dR/dT / Rt, which is the same thing as Vcc*R/(R+Rt)^2 * dR/dT.
dV_dT_ntc <- abs(ch_ntc_op$excitation_v * ch_ntc_op$divider_r_fixed_ohm
                / (ch_ntc_op$divider_r_fixed_ohm + r_ntc_op)^2 * dR_dT_ntc)

# dR/dT for the Pt1000: the CVD forward model, straightforwardly a function
# of T already, differenced the same way for a consistent method.
dR_dT_pt <- central_diff(function(t) r_pt_from_t(t, R0_FIT, A_FIT), T_OP, 1e-4)
dV_dT_pt <- abs(ch_pt_op$conditioning$gain * ch_pt_op$excitation_v
               * ch_pt_op$bridge_r_ohm
               / (r_pt_op + ch_pt_op$bridge_r_ohm)^2 * dR_dT_pt)

s1_noise <- read_run(S1_NOISE_RUN)
PIN_S1_SHORTED <- 16
sigma_floor <- sd(s1_noise$data$volts[s1_noise$data$pin == PIN_S1_SHORTED])

resolution_ntc_mK <- sigma_floor / dV_dT_ntc * 1e3
resolution_pt_mK <- sigma_floor / dV_dT_pt * 1e3
```

```{r}
#| label: out-sensitivity
#| echo: false
#| output: asis
cat(sprintf('
| Sensor | $\\mathrm{d}V/\\mathrm{d}T$ at %.0f degC (mV/K) | S1 noise floor $\\sigma$ (mV) | Resolution $\\delta T_\\mathrm{res}$ (mK) |
|---|---|---|---|
| NTC (divider) | %.3f | %.4f | %.1f |
| Pt1000 (bridge) | %.3f | %.4f | %.1f |

: Sensitivity and resolution, evaluated at the %.0f degC bath point {#tbl-t1-sensitivity}
', T_OP, dV_dT_ntc * 1e3, sigma_floor * 1e3, resolution_ntc_mK,
   dV_dT_pt * 1e3, sigma_floor * 1e3, resolution_pt_mK, T_OP))
```

```{r}
#| label: out-sensitivity-text
#| echo: false
#| output: asis
winner <- if (resolution_ntc_mK < resolution_pt_mK) "NTC" else "Pt1000"
cat(sprintf('
At %.0f degC, **%s** gives the better resolution against S1\'s own
noise floor, even though the two sensors\' raw sensitivity in mV/K may not
rank the same way — the amplifier ahead of the Pt1000 changes both the
sensitivity and, by the assumption stated in [the handbook](handbook.qmd#sec-t1-sensitivity),
contributes no separately characterized noise of its own here, so the
comparison is fair only to the extent that assumption holds.
', T_OP, winner))
```

```{r}
#| label: fig-t1-sensitivity
#| echo: false
#| fig-width: 7.0
#| fig-height: 3.6
#| fig-cap: "Empirical sensitivity between consecutive bath points, both sensors, across the calibration sweep. The NTC's sensitivity varies markedly with temperature; the bridge-plus-amplifier stage is comparatively flat."
# Empirical dV/dT between consecutive bath points -- a finite difference of
# the calibration table itself, needing no fitted model to invert.
t_mid_ntc <- (ntc$bath_C[-1] + ntc$bath_C[-nrow(ntc)]) / 2
dvdt_ntc <- diff(ntc$mean_V) / diff(ntc$bath_C)
t_mid_pt <- (pt$bath_C[-1] + pt$bath_C[-nrow(pt)]) / 2
dvdt_pt <- diff(pt$mean_V) / diff(pt$bath_C)

sens <- rbind(
  data.frame(sensor = "NTC (divider)", t_mid = t_mid_ntc, dvdt = dvdt_ntc * 1e3),
  data.frame(sensor = "Pt1000 (bridge)", t_mid = t_mid_pt, dvdt = dvdt_pt * 1e3)
)

ggplot(sens, aes(t_mid, dvdt, colour = sensor, shape = sensor)) +
  geom_line(linewidth = 0.4) +
  geom_point(size = 2.2) +
  labs(x = "Bath temperature (degC)", y = "Empirical dV/dT (mV/K)",
       colour = NULL, shape = NULL)
```

## Self-heating

Both sensors were recorded at the `t50` bath point's temperature twice: once
at the low excitation used throughout calibration, once at a higher
excitation in `selfheat-high`. Reading both through the already-fitted
curves and solving [the self-heating relation](handbook.qmd#eq-t1-selfheat-fit)
recovers $\delta$ for each sensor.

```{r}
#| label: selfheat
ch_ntc_low <- ch_ntc_op
ch_ntc_high <- channel_for_pin(selfheat, PIN_NTC)
ch_pt_low <- ch_pt_op
ch_pt_high <- channel_for_pin(selfheat, PIN_PT)

# The low-excitation point is the already-corrected voltage and resistance
# from the calibration table -- not recomputed here either.
v_ntc_low <- ntc$mean_V[ntc$run == op$meta$run]
r_ntc_low <- r_ntc_op
r_pt_low <- r_pt_op
v_ntc_high <- s1_correct(mean(selfheat$data$volts[selfheat$data$pin == PIN_NTC]))
v_pt_high <- s1_correct(mean(selfheat$data$volts[selfheat$data$pin == PIN_PT]))

r_ntc_high <- r_ntc_from_v(v_ntc_high, ch_ntc_high$divider_r_fixed_ohm, ch_ntc_high$excitation_v)
t_ntc_low <- t_ntc_from_r(r_ntc_low, coef_sh)
t_ntc_high <- t_ntc_from_r(r_ntc_high, coef_sh)
p_ntc_low <- (ch_ntc_low$excitation_v / (ch_ntc_low$divider_r_fixed_ohm + r_ntc_low))^2 * r_ntc_low
p_ntc_high <- (ch_ntc_high$excitation_v / (ch_ntc_high$divider_r_fixed_ohm + r_ntc_high))^2 * r_ntc_high
dT_ntc <- t_ntc_high - t_ntc_low
delta_ntc_mw_k <- (p_ntc_high - p_ntc_low) * 1e3 / dT_ntc

r_pt_high <- r_pt_from_v(v_pt_high, ch_pt_high$bridge_r_ohm, ch_pt_high$excitation_v,
                         ch_pt_high$conditioning$gain, ch_pt_high$conditioning$offset)
t_pt_low <- t_pt_from_r(r_pt_low, R0_FIT, A_FIT)
t_pt_high <- t_pt_from_r(r_pt_high, R0_FIT, A_FIT)
p_pt_low <- (ch_pt_low$excitation_v / (ch_pt_low$bridge_r_ohm + r_pt_low))^2 * r_pt_low
p_pt_high <- (ch_pt_high$excitation_v / (ch_pt_high$bridge_r_ohm + r_pt_high))^2 * r_pt_high
dT_pt <- t_pt_high - t_pt_low
delta_pt_mw_k <- (p_pt_high - p_pt_low) * 1e3 / dT_pt
```

```{r}
#| label: out-selfheat
#| echo: false
#| output: asis
cat(sprintf('
| Sensor | $V$ low / high (V) | $P$ low / high (mW) | $\\Delta T_\\mathrm{sh}$ (K) | $\\delta$ (mW/K) |
|---|---|---|---|---|
| NTC | %.2f / %.2f | %.4f / %.4f | %.4f | %.2f |
| Pt1000 | %.2f / %.2f | %.4f / %.4f | %.4f | %.2f |

: Self-heating shift and recovered dissipation constant, at the %.0f degC bath point {#tbl-t1-selfheat}
', ch_ntc_low$excitation_v, ch_ntc_high$excitation_v, p_ntc_low * 1e3,
   p_ntc_high * 1e3, dT_ntc, delta_ntc_mw_k,
   ch_pt_low$excitation_v, ch_pt_high$excitation_v, p_pt_low * 1e3,
   p_pt_high * 1e3, dT_pt, delta_pt_mw_k, T_OP))
```

```{r}
#| label: out-selfheat-text
#| echo: false
#| output: asis
cat(sprintf('
Raising the excitation shifted the NTC\'s apparent reading by
%.0f mK and the Pt1000\'s by %.0f mK — both well
above either fit\'s own residual scatter of a few millikelvin, and both large
enough to see clearly in a single %.0f degC reading, not only in a
200-sample mean. Neither shift was corrected for; both enter the budget below
as a bound on what a different excitation, or a different medium\'s
dissipation constant, could cost.
', dT_ntc * 1e3, dT_pt * 1e3, T_OP))
```

## Uncertainty budget

The measurand is the NTC's calibrated temperature at the `t50` bath point,
$\hat T = f(V)$ with $f$ the
composition of the divider inverse and the Steinhart-Hart curve, and the
[local sensitivity coefficient](handbook.qmd#eq-t1-budget-sens) taken by
central difference.

```{r}
#| label: budget
t_ntc_from_v_full <- function(v) {
  r <- r_ntc_from_v(v, ch_ntc_op$divider_r_fixed_ohm, ch_ntc_op$excitation_v)
  t_ntc_from_r(r, coef_sh)
}

c_v <- central_diff(t_ntc_from_v_full, v_ntc_low, 1e-6)

# Corrected voltage's standard deviation: the affine correction scales it by
# 1/B_S1, exactly as it scales the mean.
sigma_v_op <- sd(op$data$volts[op$data$pin == PIN_NTC]) / abs(B_S1)
n_op <- sum(op$data$pin == PIN_NTC)

enob_here <- enob(FSR, sigma_floor)
q_eff <- FSR / 2^enob_here

# TODO (Paolo): half-width, from the multi-point bath-uniformity check of
# thermal-rig.qmd sec-therm-bath -- not yet measured on any bench.
BATH_UNIFORMITY_K <- 0.05

components <- list(
  list(quantity = "Reference thermometer", value = T_OP, unit = "degC",
       distribution = "normal", half_width = ch_ntc_op$reference_u,
       coverage = 1.0, sensitivity = 1.0),
  list(quantity = "Fit residual (NTC)", value = 0, unit = "degC",
       distribution = "rectangular",
       half_width = max(abs(ntc$residual_K)), sensitivity = 1.0),
  list(quantity = "Effective ADC quantization", value = 0, unit = "degC",
       distribution = "rectangular", half_width = q_eff / 2, sensitivity = c_v),
  list(quantity = "Repeatability of the mean", value = v_ntc_low, unit = "degC",
       distribution = "type-a", std = sigma_v_op / sqrt(n_op), sensitivity = c_v),
  list(quantity = "Self-heating (uncorrected)", value = 0, unit = "degC",
       distribution = "rectangular", half_width = abs(dT_ntc), sensitivity = 1.0),
  list(quantity = "Bath spatial non-uniformity", value = 0, unit = "degC",
       distribution = "rectangular", half_width = BATH_UNIFORMITY_K,
       sensitivity = 1.0)
)
budget <- budget_table(components, k = 2, unit = "degC")
```

```{r}
#| label: out-budget
#| echo: false
#| output: asis
cat(budget_to_markdown(
  budget,
  caption = sprintf("Uncertainty budget for the NTC's calibrated temperature at the %.0f degC bath point", T_OP),
  label = "tbl-t1-budget"))
```

```{r}
#| label: out-budget-text
#| echo: false
#| output: asis
dominant <- budget$table[which.max(budget$table$index_pct), ]
cat(sprintf('
Combined standard uncertainty $u_c = %.1f$ mK, expanded
$U = %.1f$ mK at $k = 2$. The dominant component is
**%s** at %.0f %% of the combined
variance.
', budget$u_c * 1e3, budget$U * 1e3, dominant$quantity, dominant$index_pct))
```

## What to report

1. @tbl-t1-ntc-cal and @fig-t1-ntc-cal: the NTC's Steinhart-Hart fit, with
   $A$, $B$, $C$ and the residuals in millikelvin.
2. @tbl-t1-pt-cal and @fig-t1-pt-cal: the Pt1000's Callendar-Van Dusen fit,
   with $R_0$, the recovered $A$ against the fixed IEC 60751 value, and the
   residuals.
3. @tbl-t1-sensitivity and @fig-t1-sensitivity: sensitivity and resolution
   for both sensors, and which one wins against S1's own noise floor.
4. @tbl-t1-selfheat: the self-heating shift and the recovered dissipation
   constant for both sensors.
5. @tbl-t1-budget: the budget, the dominant component, and what you would
   change first to reduce it.
