---
title: "T2 — Analysis template (R)"
subtitle: "Dynamic response of temperature sensors"
description: "Time constants, convection coefficients, the ramp error and a second-order probe, in R"
engine: knitr
toc: true
toc-depth: 3
code-fold: false
---

This template runs end to end on the reference data shipped with the
experiment. To analyse your own measurements, replace the files in `data/`
and change the paths in the first chunk — nothing else.

The Python template [`analysis-py.qmd`](analysis-py.qmd) has the same
sections, the same figure labels and the same output tables, and produces the
same numbers from the same data.

::: {.callout-note collapse="true"}
## Two things this template does deliberately

**It refits T1's calibration rather than hard-coding it.** Volts become
temperature through the Steinhart–Hart and Callendar–Van Dusen curves fitted
from T1's own recordings, read from `../T1-temperature-static/data/`. A
calibration is an input quantity with an uncertainty, not a constant.

**Every fit is written out.** The nonlinear fits are explicit damped
Gauss–Newton loops rather than `scipy.optimize.curve_fit` against R's `nls`,
which choose their own convergence criteria and do not reach the same
parameters on the same data.
:::

## Setup

```{r}
#| label: setup
#| message: false
library(ggplot2)
library(patchwork)

source("../../shared/R/labtools.R")

theme_set(theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.25)))

STEP_RUNS <- c("step-still-air", "step-forced-air", "step-still-water",
               "step-stirred-water")
SHEATHED_RUN <- "step-sheathed"
RAMP_RUN <- "ramp-stirred-water"
PRIMARY_MEDIUM <- "stirred-water"     # the budget's medium

T1_DATA <- "../T1-temperature-static/data"
B_IEC <- -5.775e-7                    # IEC 60751, fixed
MAX_FIT_ITERATIONS <- 200

# rho*c*V/A for the bare probe, in J/(m^2 K). An absolute h rests entirely on
# this number, which this experiment does not measure.
# TODO (Paolo): measure or estimate for the probe actually used.
RHO_C_V_OVER_A <- 4.0e3
```

```{r}
#| label: helpers
# Inverts the divider for the NTC's resistance, as in T1.
r_ntc_from_v <- function(v, r_fixed, vcc) vcc * r_fixed / v - r_fixed

# Steinhart-Hart from a fitted (A, B, C), degrees Celsius.
t_ntc_from_r <- function(r, coef) {
  ln_r <- log(r)
  1 / (coef[1] + coef[2] * ln_r + coef[3] * ln_r^3) - 273.15
}

# Inverts the bridge for the Pt1000's resistance, as in T1.
r_pt_from_v <- function(v, r_bridge, v_exc, gain, offset) {
  deflection <- (v - offset) / (gain * v_exc)
  r_bridge * (0.5 + deflection) / (0.5 - deflection)
}

# Callendar-Van Dusen, positive root of the quadratic in t.
t_pt_from_r <- function(r, r0, a, b = B_IEC) {
  cc <- 1 - r / r0
  (-a + sqrt(a^2 - 4 * b * cc)) / (2 * b)
}

# Damped Gauss-Newton on an arbitrary model. `rj(p)` returns list(r, j) with r
# the residual vector and j its Jacobian. The step is halved until the sum of
# squares does not increase, which is deterministic and therefore identical in
# both languages; an undamped step diverges on the slower media, where the
# baseline pins T_0 much harder than the tail pins tau.
gauss_newton <- function(rj, p0, n_obs) {
  p <- as.numeric(p0)
  for (i in seq_len(MAX_FIT_ITERATIONS)) {
    out <- rj(p)
    sse <- sum(out$r^2)
    step <- as.numeric(solve(crossprod(out$j), crossprod(out$j, out$r)))
    lam <- 1
    while (lam > 1e-8) {
      if (sum(rj(p + lam * step)$r^2) <= sse) break
      lam <- lam / 2
    }
    p <- p + lam * step
    if (max(abs(lam * step) / pmax(abs(p), 1e-12)) < 1e-12) break
  }
  out <- rj(p)
  dof <- max(n_obs - length(p), 1)
  s <- sqrt(sum(out$r^2) / dof)
  cov <- solve(crossprod(out$j)) * s^2
  list(p = p, u = sqrt(diag(cov)), s = s, residual = out$r)
}

# Fits the first-order step response, returning all three parameters.
fit_first_order <- function(t, y) {
  rj <- function(p) {
    t_inf <- p[1]; t_0 <- p[2]; tau <- p[3]
    e <- exp(-t / tau)
    list(r = y - (t_inf + (t_0 - t_inf) * e),
         j = cbind(1 - e, e, (t_0 - t_inf) * e * t / tau^2))
  }
  p0 <- c(y[length(y)], y[1], max((t[length(t)] - t[1]) / 5, 1e-3))
  fit <- gauss_newton(rj, p0, length(y))
  list(T_inf = fit$p[1], T_0 = fit$p[2], tau = fit$p[3],
       u_tau = fit$u[3], u_T_inf = fit$u[1], s_res = fit$s,
       residual = fit$residual, t = t, y = y)
}

# The log-linear estimator, for comparison with the nonlinear fit.
fit_linearised <- function(t, y, t_inf) {
  z <- abs(y - t_inf)
  keep <- z > 1e-9
  x <- t[keep]; w <- log(z[keep])
  sxx <- sum((x - mean(x))^2)
  slope <- sum((x - mean(x)) * (w - mean(w))) / sxx
  -1 / slope
}

# Fits the two-lump second-order response.
fit_second_order <- function(t, y) {
  shape <- function(t1, t2) {
    (t1 * exp(-t / t1) - t2 * exp(-t / t2)) / (t1 - t2)
  }
  rj <- function(p) {
    t_inf <- p[1]; t_0 <- p[2]; t1 <- p[3]; t2 <- p[4]
    g <- shape(t1, t2)
    # Numerical columns for the two time constants: their closed forms are
    # long and the step size below is far smaller than the fit's own
    # precision, so both languages get the same derivative.
    eps <- 1e-6
    d1 <- (shape(t1 + eps, t2) - shape(t1 - eps, t2)) / (2 * eps)
    d2 <- (shape(t1, t2 + eps) - shape(t1, t2 - eps)) / (2 * eps)
    list(r = y - (t_inf + (t_0 - t_inf) * g),
         j = cbind(1 - g, g, (t_0 - t_inf) * d1, (t_0 - t_inf) * d2))
  }
  span <- t[length(t)] - t[1]
  p0 <- c(y[length(y)], y[1], span / 20, span / 6)
  fit <- gauss_newton(rj, p0, length(y))
  list(T_inf = fit$p[1], T_0 = fit$p[2], tau1 = fit$p[3], tau2 = fit$p[4],
       u_tau1 = fit$u[3], u_tau2 = fit$u[4], s_res = fit$s,
       residual = fit$residual)
}

ols <- function(x, y) {
  sxx <- sum((x - mean(x))^2)
  slope <- sum((x - mean(x)) * (y - mean(y))) / sxx
  intercept <- mean(y) - slope * mean(x)
  resid <- y - (intercept + slope * x)
  s <- sqrt(sum(resid^2) / max(length(x) - 2, 1))
  list(slope = slope, intercept = intercept, u_slope = s / sqrt(sxx), s_res = s)
}
```

## Refitting T1's calibration

```{r}
#| label: t1-calibration
t1_runs <- lapply(sort(Sys.glob(file.path(T1_DATA, "cal-*.csv"))), read_run)

ntc_r <- c(); pt_r <- c(); t_ref <- c()
for (run in t1_runs) {
  for (ch in run$meta$channels) {
    v <- mean(run$data$volts[run$data$pin == ch$pin])
    if (!is.null(ch$divider_r_fixed_ohm)) {
      ntc_r <- c(ntc_r, r_ntc_from_v(v, ch$divider_r_fixed_ohm, ch$excitation_v))
    } else {
      pt_r <- c(pt_r, r_pt_from_v(v, ch$bridge_r_ohm, ch$excitation_v,
                                  ch$conditioning$gain, ch$conditioning$offset))
    }
  }
  t_ref <- c(t_ref, run$meta$channels[[1]]$reference_value)
}

# Steinhart-Hart: linear in (1, ln R, (ln R)^3) against 1/T in kelvin.
ln_r <- log(ntc_r)
design <- cbind(rep(1, length(ln_r)), ln_r, ln_r^3)
sh_coef <- as.numeric(solve(crossprod(design),
                            crossprod(design, 1 / (t_ref + 273.15))))
ntc_resid_k <- t_ntc_from_r(ntc_r, sh_coef) - t_ref

# Callendar-Van Dusen with B fixed: linear in (R0, R0*A) against (1+Bt^2, t).
design_pt <- cbind(1 + B_IEC * t_ref^2, t_ref)
pt_x <- as.numeric(solve(crossprod(design_pt), crossprod(design_pt, pt_r)))
R0 <- pt_x[1]; A_CVD <- pt_x[2] / pt_x[1]
pt_resid_k <- t_pt_from_r(pt_r, R0, A_CVD) - t_ref
```

```{r}
#| label: load
steps <- lapply(STEP_RUNS, function(n) read_run(file.path("data", paste0(n, ".csv"))))
names(steps) <- STEP_RUNS
sheathed <- read_run(file.path("data", paste0(SHEATHED_RUN, ".csv")))
ramp <- read_run(file.path("data", paste0(RAMP_RUN, ".csv")))
board <- steps[[1]]$meta$board
LSB <- lsb_volts(steps[[1]]$meta)
```

```{r}
#| label: banner
#| echo: false
#| output: asis
cat(synthetic_banner(steps[[1]]$meta))
```

```{r}
#| label: out-calibration
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| T1 calibration points used | %d |
| Steinhart–Hart $A$ | %.6e |
| Steinhart–Hart $B$ | %.6e |
| Steinhart–Hart $C$ | %.6e |
| NTC fit residual (RMS) | %.1f mK |
| Pt1000 $R_0$ | %.3f Ω |
| Pt1000 $A$ | %.6e K⁻¹ |
| Pt1000 fit residual (RMS) | %.1f mK |

: T1\'s calibration, refitted here from T1\'s own recordings {#tbl-t2-calibration}
', length(t1_runs), sh_coef[1], sh_coef[2], sh_coef[3],
   sqrt(mean(ntc_resid_k^2)) * 1e3, R0, A_CVD, sqrt(mean(pt_resid_k^2)) * 1e3))
```

## Data quality

```{r}
#| label: quality
#| echo: false
#| output: asis
all_runs <- c(steps, list(sheathed), list(ramp))
bad <- sum(vapply(all_runs, function(r) length(validate_meta(r$meta)) > 0, logical(1)))
lossy <- character(0)
for (r in all_runs) {
  st <- r$meta$stream_stats
  counters <- st[setdiff(names(st), "records")]
  if (any(vapply(counters, function(v) is.numeric(v) && v > 0, logical(1)))) {
    lossy <- c(lossy, r$meta$run)
  }
}
cat(sprintf('
%d runs read: %d plunges, one sheathed probe and one ramp.
%s
%s
', length(all_runs), length(steps),
   if (bad == 0) "All sidecars are complete."
   else sprintf("%d sidecar(s) have missing fields.", bad),
   if (length(lossy) == 0) "No run reports a lost record."
   else paste0("Runs with losses: ", paste(lossy, collapse = ", "))))
```

## Converting to temperature

```{r}
#| label: to-temperature
# Volts to degrees Celsius, through T1's refitted curve for that sensor.
temperature_of <- function(run) {
  ch <- run$meta$channels[[1]]
  d <- run$data[run$data$pin == ch$pin, ]
  t_s <- d$t_us / 1e6
  t_s <- t_s - t_s[1]
  v <- d$volts
  temp <- if (!is.null(ch$divider_r_fixed_ohm)) {
    t_ntc_from_r(r_ntc_from_v(v, ch$divider_r_fixed_ohm, ch$excitation_v), sh_coef)
  } else {
    t_pt_from_r(r_pt_from_v(v, ch$bridge_r_ohm, ch$excitation_v,
                            ch$conditioning$gain, ch$conditioning$offset),
                R0, A_CVD)
  }
  list(t = t_s, temp = temp, ch = ch)
}

# The record from the plunge onward, with time measured from it.
after_plunge <- function(run) {
  o <- temperature_of(run)
  t0 <- o$ch$plunge_time_s
  keep <- o$t >= t0
  list(t_rel = o$t[keep] - t0, temp = o$temp[keep],
       t_full = o$t, temp_full = o$temp, ch = o$ch)
}
```

## Time constants

```{r}
#| label: step-fits
fits <- list()
rows <- list()
for (name in STEP_RUNS) {
  a <- after_plunge(steps[[name]])
  fit <- fit_first_order(a$t_rel, a$temp)
  tau_lin <- fit_linearised(a$t_rel, a$temp, fit$T_inf)
  fits[[name]] <- list(fit = fit, t_rel = a$t_rel, temp = a$temp, ch = a$ch)
  rows[[name]] <- data.frame(
    medium = a$ch$medium, n = length(a$temp),
    T_0 = fit$T_0, T_inf = fit$T_inf,
    tau_s = fit$tau, u_tau_s = fit$u_tau,
    tau_lin_s = tau_lin,
    lin_diff_pct = (tau_lin / fit$tau - 1) * 100,
    s_res_mK = fit$s_res * 1e3,
    span_over_tau = (a$t_rel[length(a$t_rel)] - a$t_rel[1]) / fit$tau)
}
steps_df <- do.call(rbind, rows)
steps_df <- steps_df[order(-steps_df$tau_s), ]
rownames(steps_df) <- NULL
```

```{r}
#| label: out-steps
#| echo: false
#| output: asis
rows <- sprintf("| %s | %d | %.2f | %.2f | %.3f | %.1f | %.1f | %.1f |",
                steps_df$medium, steps_df$n, steps_df$T_0, steps_df$T_inf,
                steps_df$tau_s, steps_df$u_tau_s * 1e3,
                steps_df$span_over_tau, steps_df$s_res_mK)
cat(sprintf('
| Medium | $n$ | $T_0$ [°C] | $T_\\infty$ [°C] | $\\tau$ [s] | $u(\\tau)$ [ms] | Record / $\\tau$ | Residual [mK] |
|---|---|---|---|---|---|---|---|
%s

: Time constants from the nonlinear fit of the [step response](handbook.qmd#eq-t2-step) {#tbl-t2-tau}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: fig-t2-steps
#| echo: false
#| fig-width: 7.8
#| fig-height: 3.6
#| fig-cap: "The four plunges, each on its own time axis scaled by its own fitted time constant. On this scaling a first-order response is the same curve in every medium — which is the claim the experiment is testing."
raw <- do.call(rbind, lapply(STEP_RUNS, function(n) {
  f <- fits[[n]]
  data.frame(medium = f$ch$medium, t = f$t_rel, temp = f$temp,
             norm = (f$temp - f$fit$T_inf) / (f$fit$T_0 - f$fit$T_inf),
             t_over_tau = f$t_rel / f$fit$tau)
}))

p1 <- ggplot(raw[raw$t > 0, ], aes(t, temp, colour = medium)) +
  geom_line(linewidth = 0.3) +
  scale_x_log10() +
  labs(x = "Time since plunge [s]", y = "Temperature [°C]", colour = NULL) +
  theme(legend.position = "bottom", legend.text = element_text(size = 7))

grid <- data.frame(x = seq(0, 5, length.out = 200))
grid$y <- exp(-grid$x)
p2 <- ggplot(raw[raw$t_over_tau <= 5, ], aes(t_over_tau, norm, colour = medium)) +
  geom_line(linewidth = 0.3, show.legend = FALSE) +
  geom_line(data = grid, aes(x, y), inherit.aes = FALSE,
            linetype = "dashed", linewidth = 0.4) +
  labs(x = "Time since plunge, in units of tau", y = "Normalised approach")

p1 | p2
```

```{r}
#| label: out-estimators
#| echo: false
#| output: asis
rows <- sprintf("| %s | %.3f | %.3f | %+.2f |",
                steps_df$medium, steps_df$tau_s, steps_df$tau_lin_s,
                steps_df$lin_diff_pct)
cat(sprintf('
| Medium | $\\tau$ nonlinear [s] | $\\tau$ linearised [s] | Difference |
|---|---|---|---|
%s

: The two estimators of the handbook do not agree {#tbl-t2-estimators}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: out-estimators-text
#| echo: false
#| output: asis
worst <- steps_df[which.max(abs(steps_df$lin_diff_pct)), ]
cat(sprintf('
The linearised estimator differs from the nonlinear one by up to
%.1f %% (in %s), which is far larger
than either fit\'s own standard error. The log transform gives the tail — where
$T-T_\\infty$ is a few millikelvin and almost entirely noise — the same weight
as the first, most informative, part of the curve. Report the nonlinear
figure, and quote the linearised one only to show the difference.
', abs(worst$lin_diff_pct), worst$medium))
```

## Convection coefficients

```{r}
#| label: convection
fastest <- steps_df[nrow(steps_df), ]      # smallest tau
conv <- steps_df[, c("medium", "tau_s", "u_tau_s")]
conv$h_ratio <- fastest$tau_s / conv$tau_s
conv$h_absolute <- RHO_C_V_OVER_A / conv$tau_s
```

```{r}
#| label: out-convection
#| echo: false
#| output: asis
rows <- sprintf("| %s | %.3f | %.3f | %.0f |",
                conv$medium, conv$tau_s, conv$h_ratio, conv$h_absolute)
cat(sprintf('
| Medium | $\\tau$ [s] | $h/h_\\mathrm{%s}$ | $h$ [W m⁻² K⁻¹] |
|---|---|---|---|
%s

: Convection, as ratios (measured) and as absolute values (estimated) {#tbl-t2-convection}
', fastest$medium, paste(rows, collapse = "\n")))
```

```{r}
#| label: out-convection-text
#| echo: false
#| output: asis
slowest <- conv[1, ]
cat(sprintf('
The ratio column is a **measurement**: the [ratio form](handbook.qmd#eq-t2-h-ratio) needs no probe geometry,
because $\\rho cV/A$ is identical in every row and cancels. Stirred water
removes heat %.0f times faster than
%s does, and that number stands on its own.

The absolute column is an **estimate**. It rests entirely on the assumed
$\\rho cV/A = %.0f$ J m⁻² K⁻¹, which this experiment does not
measure and which is marked TODO in the template. Quote the ratios as results
and the absolute values with that assumption attached, or not at all.
', 1 / slowest$h_ratio, slowest$medium, RHO_C_V_OVER_A))
```

## The ramp error

```{r}
#| label: ramp
o_ramp <- temperature_of(ramp)
t_ramp <- o_ramp$t; temp_ramp <- o_ramp$temp; ch_ramp <- o_ramp$ch
rate <- ch_ramp$ramp_rate_c_per_s
t_start <- ch_ramp$ramp_start_time_s
t0_ramp <- ch_ramp$initial_temperature_c

# The true bath temperature the sensor is following.
true_temp <- ifelse(t_ramp < t_start, t0_ramp,
                    t0_ramp + rate * (t_ramp - t_start))
error <- temp_ramp - true_temp

tau_stirred <- steps_df$tau_s[steps_df$medium == PRIMARY_MEDIUM]
u_tau_stirred <- steps_df$u_tau_s[steps_df$medium == PRIMARY_MEDIUM]
predicted <- -tau_stirred * rate

# Measure the settled error only where the transient has genuinely died. The
# procedure's own rule -- five time constants -- leaves exp(-5) = 0.7 % of the
# initial lag, well under the scatter of the mean below; stopping at four
# leaves 1.8 %, which is the same size as the whole discrepancy and would be
# read as a disagreement with theory rather than as an impatient window.
SETTLE_TAUS <- 5
settled <- t_ramp > t_start + SETTLE_TAUS * tau_stirred
measured <- mean(error[settled])
u_measured <- sd(error[settled]) / sqrt(sum(settled))
```

```{r}
#| label: fig-t2-ramp
#| echo: false
#| fig-width: 7.0
#| fig-height: 4.6
#| fig-cap: "A sensor following a ramp. Top: the bath and the sensor. Bottom: their difference, which settles to a constant lag rather than decaying to zero. The dashed line is the prediction from the step test, with no free parameter."
traces <- rbind(
  data.frame(t = t_ramp, y = true_temp, what = "bath"),
  data.frame(t = t_ramp, y = temp_ramp, what = "sensor"))
p_t <- ggplot(traces, aes(t, y, colour = what, linetype = what)) +
  geom_line(linewidth = 0.4) +
  labs(x = NULL, y = "Temperature [°C]", colour = NULL, linetype = NULL) +
  theme(legend.position = "top")

p_e <- ggplot(data.frame(t = t_ramp, e = error), aes(t, e)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.3) +
  geom_hline(yintercept = predicted, linewidth = 0.4, linetype = "dashed") +
  labs(x = "Time [s]", y = "Sensor − bath [K]",
       subtitle = sprintf("dashed: -tau*rate = %.3f K", predicted))

p_t / p_e + plot_layout(heights = c(2, 1))
```

```{r}
#| label: out-ramp
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| Ramp rate $\\dot{T}$ | %.3f K s⁻¹ |
| $\\tau$ from the step test (%s) | %.3f ± %.3f s |
| Predicted steady error $-\\tau\\dot{T}$ | %.3f K |
| Measured steady error | %.3f ± %.3f K |
| Difference | %+.3f K |

: The ramp error, predicted from the step test and measured {#tbl-t2-ramp}
', rate, PRIMARY_MEDIUM, tau_stirred, u_tau_stirred, predicted,
   measured, u_measured, measured - predicted))
```

```{r}
#| label: out-ramp-text
#| echo: false
#| output: asis
agree <- abs(measured - predicted) < 2 * sqrt(u_measured^2 + (u_tau_stirred * rate)^2)
cat(sprintf('
The prediction uses **no free parameter**: $\\tau$ comes from a different
recording, in a different experiment section, and $\\dot{T}$ from the
controller. The two
%s.

Note what kind of error this is. It does not decay, it does not average down,
and a static calibration cannot see it — the sensor is perfectly calibrated and
still reads %.2f K low for as long as the ramp continues.
', if (agree) "agree within their combined uncertainty"
   else "do not agree within their combined uncertainty, which is itself a result worth explaining",
   abs(measured)))
```

## A second-order probe

```{r}
#| label: sheathed
a_sh <- after_plunge(sheathed)
t_sh <- a_sh$t_rel; temp_sh <- a_sh$temp
fit1 <- fit_first_order(t_sh, temp_sh)
fit2 <- fit_second_order(t_sh, temp_sh)
```

```{r}
#| label: fig-t2-sheathed
#| echo: false
#| fig-width: 7.0
#| fig-height: 4.8
#| fig-cap: "The sheathed probe. Top: the first seconds, where a first-order model must have its maximum slope at the plunge and this response does not. Bottom: the residual of the first-order fit, whose arc — negative, positive, negative — is the diagnostic, not its magnitude."
early <- t_sh <= min(6 * fit1$tau, t_sh[length(t_sh)])
model1 <- fit1$T_inf + (fit1$T_0 - fit1$T_inf) * exp(-t_sh / fit1$tau)
g <- (fit2$tau1 * exp(-t_sh / fit2$tau1) - fit2$tau2 * exp(-t_sh / fit2$tau2)) /
  (fit2$tau1 - fit2$tau2)
model2 <- fit2$T_inf + (fit2$T_0 - fit2$T_inf) * g

p_fit <- ggplot() +
  geom_point(data = data.frame(t = t_sh[early], y = temp_sh[early]),
             aes(t, y), size = 0.6) +
  geom_line(data = data.frame(t = t_sh[early], y = model1[early], w = "first order"),
            aes(t, y, colour = w), linewidth = 0.45) +
  geom_line(data = data.frame(t = t_sh[early], y = model2[early], w = "second order"),
            aes(t, y, colour = w), linewidth = 0.45) +
  labs(x = NULL, y = "Temperature [°C]", colour = NULL) +
  theme(legend.position = "top")

res <- rbind(
  data.frame(t = t_sh[early], r = fit1$residual[early] * 1e3, w = "first order"),
  data.frame(t = t_sh[early], r = fit2$residual[early] * 1e3, w = "second order"))
p_res <- ggplot(res, aes(t, r, colour = w)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.3, show.legend = FALSE) +
  labs(x = "Time since plunge [s]", y = "Residual [mK]")

p_fit / p_res + plot_layout(heights = c(2, 1))
```

```{r}
#| label: out-sheathed
#| echo: false
#| output: asis
cat(sprintf('
| Model | Parameters | Residual RMS |
|---|---|---|
| First order | $\\tau = %.3f \\pm %.3f$ s | %.1f mK |
| Second order | $\\tau_1 = %.3f \\pm %.3f$ s, $\\tau_2 = %.3f \\pm %.3f$ s | %.1f mK |

: One lump against two, on the sheathed probe {#tbl-t2-sheathed}
', fit1$tau, fit1$u_tau, fit1$s_res * 1e3,
   fit2$tau1, fit2$u_tau1, fit2$tau2, fit2$u_tau2, fit2$s_res * 1e3))
```

```{r}
#| label: out-sheathed-text
#| echo: false
#| output: asis
cat(sprintf('
The second model reduces the residual by a factor of %.1f, but
the number to look at is the **shape** in @fig-t2-sheathed. The first-order
residual is not scatter: it is negative at the start, positive through the
middle and negative again, which is what an unmodelled second lag always looks
like. A single goodness-of-fit number would have hidden it, and a first-order
$\\tau$ of %.2f s would have been reported without anyone
noticing that the probe does not have one.
', fit1$s_res / fit2$s_res, fit1$tau))
```

## Uncertainty budget

The measurand is $\tau$ in stirred water, the fastest medium and therefore
the one where every systematic effect is largest.

```{r}
#| label: budget
primary <- fits[[paste0("step-", PRIMARY_MEDIUM)]]
fit_p <- primary$fit
ch_p <- primary$ch

# The transfer is not instantaneous; the bound is the measured transfer time.
# TODO (Paolo): replace with the transfer duration timed at the bench.
T_TRANSFER_S <- 0.2

# T1's calibration residual, referred to tau through the fit's own sensitivity
# dtau/dT: a temperature offset of dT shifts the fitted tau by roughly
# tau * dT / (T_0 - T_inf).
cal_resid_k <- sqrt(mean(ntc_resid_k^2))
c_cal <- fit_p$tau / abs(fit_p$T_0 - fit_p$T_inf)

# The time axis is only as good as the sampling grid (S1).
grid_stats <- timing_summary(steps[[paste0("step-", PRIMARY_MEDIUM)]]$data)$stats
grid_bound_s <- grid_stats$residual_max_abs_us / 1e6

components <- list(
  list(quantity = "Fit repeatability", value = fit_p$tau, unit = "s",
       distribution = "type-a", std = fit_p$u_tau, sensitivity = 1),
  list(quantity = "Transfer time", value = 0, unit = "s",
       distribution = "rectangular", half_width = T_TRANSFER_S / 2,
       sensitivity = 1),
  list(quantity = "T1 calibration", value = 0, unit = "s",
       distribution = "type-a", std = cal_resid_k, sensitivity = c_cal),
  list(quantity = "Sampling grid", value = 0, unit = "s",
       distribution = "rectangular", half_width = grid_bound_s, sensitivity = 1),
  list(quantity = "Bath stability", value = 0, unit = "s",
       distribution = "type-a", std = fit_p$u_T_inf, sensitivity = c_cal)
)
budget <- budget_table(components, k = 2)
```

```{r}
#| label: out-budget
#| echo: false
#| output: asis
cat(budget_to_markdown(
  budget,
  caption = sprintf("Uncertainty budget for $\\tau$ in %s", ch_p$medium),
  label = "tbl-t2-budget"))
```

```{r}
#| label: out-budget-text
#| echo: false
#| output: asis
dominant <- budget$table[which.max(budget$table$index_pct), ]
cat(sprintf('
$\\tau = %.3f$ s with an expanded uncertainty
$U = %.0f$ ms at $k = 2$, a relative uncertainty of
%.1f %%.

The dominant component is **%s** at
%.0f %% of the combined variance. If that is the
transfer time, the way to improve this measurement is a faster plunge, not a
longer record or a quieter converter — and the fit\'s own standard error,
%.0f ms, would have been a badly optimistic thing to
quote on its own.
', fit_p$tau, budget$U * 1e3, budget$U / fit_p$tau * 100,
   dominant$quantity, dominant$index_pct, fit_p$u_tau * 1e3))
```

## What to report

1. @tbl-t2-calibration: T1's calibration as you refitted it, with its residual.
2. @tbl-t2-tau and @fig-t2-steps: a time constant per medium, each with the
   medium named, and the record length in units of $\tau$.
3. @tbl-t2-estimators: the two estimators, their difference, and which you
   report.
4. @tbl-t2-convection: the ratios as measurements and any absolute $h$ with
   its assumption stated.
5. @tbl-t2-ramp and @fig-t2-ramp: the predicted and measured ramp error, and
   one sentence on why averaging does not reduce it.
6. @tbl-t2-sheathed and @fig-t2-sheathed: the residual's shape, not its size.
7. @tbl-t2-budget: the budget, the dominant component, and what you would
   change first.
