---
title: "T3 — Analysis template (R)"
subtitle: "Long-term stability and drift"
description: "Drift, diurnal cycle and the Allan-deviation optimum, in R"
engine: knitr
toc: true
toc-depth: 3
code-fold: false
---

This template runs end to end on the reference data shipped with the
experiment, so you can read the whole analysis before your own 24-hour run
finishes. To analyse your own measurement, replace the file in `data/` and
change the path in the first chunk — nothing else.

The Python template [`analysis-py.qmd`](analysis-py.qmd) has the same
sections, the same figure labels and the same output tables, and produces the
same numbers from the same data.

::: {.callout-note collapse="true"}
## Why the Allan-deviation averaging times are a fixed list

Both templates evaluate [the Allan deviation](handbook.qmd#eq-t3-allan) at the same 28 integer values of $m$,
written out once as a literal list rather than regenerated from
`logspace`/`round` in each language. `round()` is round-half-to-even in both
Python and R, so the two would very likely agree anyway — but "very likely"
is not "identical," and a table-for-table comparison needs identical, not
merely close.
:::

## Setup

```{r}
#| label: setup
#| message: false
library(ggplot2)
library(patchwork)

source("../../shared/R/labtools.R")

theme_set(theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.25)))

LOG_RUN <- "data/log-24h.csv"
DIURNAL_PERIOD_S <- 86400.0

# The averaging times at which the overlapping Allan deviation is evaluated,
# m = tau / tau0 -- see the callout above for why this is a fixed list
# instead of a freshly generated one in each language.
M_LIST <- c(1, 2, 3, 4, 6, 8, 10, 13, 18, 24, 32, 43, 57, 76, 101, 135, 181,
           241, 322, 429, 573, 764, 1020, 1362, 1817, 2425, 3237, 4320)
```

```{r}
#| label: helpers
# Overlapping Allan deviation of a rate-like series `y`, at spacing `tau0`,
# evaluated at averaging times `tau = m * tau0`. `y` is integrated once into
# a phase-like `x` (a running sum, scaled by `tau0`), exactly as [the Allan deviation](handbook.qmd#eq-t3-allan)
# is written in the handbook; the doubled, overlapping second difference of
# `x` is then formed for each `m`.
allan_deviation <- function(y, tau0, ms) {
  x <- c(0.0, cumsum(y) * tau0)
  out <- numeric(length(ms))
  for (j in seq_along(ms)) {
    m <- ms[j]
    n2m <- length(x) - 2 * m
    d <- x[(2 * m + 1):(2 * m + n2m)] - 2 * x[(m + 1):(m + n2m)] + x[1:n2m]
    out[j] <- sqrt(sum(d^2) / (2 * n2m * (m * tau0)^2))
  }
  out
}

# The sensor's own linear transduction, V = a + b * T_celsius.
volts_to_celsius <- function(volts, channel) {
  a <- channel$sensor_offset_V
  b <- channel$sensor_sensitivity_V_per_K
  (volts - a) / b
}
```

```{r}
#| label: load
run <- read_run(LOG_RUN)
board <- run$meta$board
channel <- run$meta$channels[[1]]

t_us <- run$data$t_us
t_s <- (t_us - t_us[1]) / 1e6
temp_c <- volts_to_celsius(run$data$volts, channel)

tau0 <- timing_summary(run$data)$stats$period_us / 1e6
n <- length(temp_c)
```

```{r}
#| label: banner
#| echo: false
#| output: asis
cat(synthetic_banner(run$meta))
```

## Data quality

```{r}
#| label: quality
#| echo: false
#| output: asis
problems <- validate_meta(run$meta)
stats <- run$meta$stream_stats
counters <- stats[setdiff(names(stats), "records")]
lossy <- counters[vapply(counters, function(v) is.numeric(v) && v > 0, logical(1))]
duration_h <- (t_s[n] - t_s[1]) / 3600.0
cat(sprintf('
**%s** — %d records over
%.2f h, sample spacing $\\tau_0 = %.3f$ s;
%s;
%s.
', run$meta$run, stats$records, duration_h, tau0,
   if (length(problems) == 0) "metadata complete"
   else paste0("metadata problems: ", paste(problems, collapse = "; ")),
   if (length(lossy) == 0) "no losses reported"
   else paste0("losses: ",
               paste(names(lossy), unlist(lossy), sep = " = ", collapse = ", "))))
```

## Drift and the diurnal cycle

The record is fit to [the four-parameter model](handbook.qmd#eq-t3-fit): a
mean, a linear drift, and a sinusoid fixed at exactly 24 hours.

```{r}
#| label: fit
w <- 2 * pi / DIURNAL_PERIOD_S
D <- cbind(1, t_s, cos(w * t_s), sin(w * t_s))
coef <- solve(crossprod(D), crossprod(D, temp_c))
resid <- as.vector(temp_c - D %*% coef)

drift_K_per_h <- coef[2] * 3600
diurnal_amp_K <- sqrt(coef[3]^2 + coef[4]^2)

# Standard errors of the fit, from the residuals (n - 4 degrees of freedom).
dof <- max(n - 4, 1)
s_res <- sqrt(sum(resid^2) / dof)
cov <- s_res^2 * solve(crossprod(D))
u_drift_K_per_h <- sqrt(cov[2, 2]) * 3600
```

```{r}
#| label: fig-t3-fit
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.0
#| fig-cap: "The 24-hour record (grey) with the fitted drift-plus-diurnal model (top), and the residual left over (bottom) -- what [the Allan deviation](handbook.qmd#eq-t3-allan) is computed on."
t_h <- t_s / 3600
p_fit <- ggplot() +
  geom_line(data = data.frame(t = t_h, v = temp_c), aes(t, v),
            colour = "grey60", linewidth = 0.3) +
  geom_line(data = data.frame(t = t_h, v = as.vector(D %*% coef)), aes(t, v),
            linewidth = 0.5) +
  labs(x = NULL, y = "Temperature [°C]")

p_res <- ggplot(data.frame(t = t_h, r = resid * 1e3), aes(t, r)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.2) +
  labs(x = "Time [h]", y = "Residual [mK]")

p_fit / p_res
```

```{r}
#| label: out-fit
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| Drift $\\dot x$ | $%+.3f \\pm %.3f$ mK/h |
| Diurnal amplitude $A_d$ | $%.1f$ mK |
| Residual RMS after the fit | $%.3f$ mK |
| Record length | $%.2f$ h |

: Drift and diurnal fit of [the drift-plus-diurnal model](handbook.qmd#eq-t3-fit) {#tbl-t3-fit}
', drift_K_per_h * 1e3, u_drift_K_per_h * 1e3, diurnal_amp_K * 1e3,
   s_res * 1e3, duration_h))
```

```{r}
#| label: out-fit-text
#| echo: false
#| output: asis
ratio_diurnal <- if (drift_K_per_h != 0) diurnal_amp_K / abs(drift_K_per_h * duration_h) else NaN
cat(sprintf('
The diurnal amplitude is %.0f
times the total drift accumulated over the record -- the room, attenuated by
the box, is still the largest feature, exactly as [why the box is insulated](handbook.qmd#sec-t3-why) says it should
be. What the Allan deviation below characterizes is the residual left after
removing both.
', ratio_diurnal))
```

## The Allan deviation

[the Allan deviation](handbook.qmd#eq-t3-allan) is evaluated on the detrended residual at the fixed averaging
times of `M_LIST`.

```{r}
#| label: allan
tau <- M_LIST * tau0
adev <- allan_deviation(resid, tau0, M_LIST)

# The *first* local minimum, not the global one: fitting and removing a
# deterministic trend over the whole record (the drift-plus-diurnal model
# above) systematically suppresses the residual's own variance at the
# longest averaging times -- an edge effect of detrending, not a second,
# better optimum. Scanning from short tau upward finds the genuine
# white-noise/random-walk crossover instead.
i_opt <- NA
for (i in 2:(length(adev) - 1)) {
  if (adev[i] <= adev[i - 1] && adev[i] <= adev[i + 1]) {
    i_opt <- i
    break
  }
}
if (is.na(i_opt)) i_opt <- which.min(adev)
tau_opt <- tau[i_opt]
sigma_opt <- adev[i_opt]
```

```{r}
#| label: fig-t3-allan
#| echo: false
#| fig-width: 7.0
#| fig-height: 4.4
#| fig-cap: "Overlapping Allan deviation of the detrended residual. Dashed lines: the two slopes of [the slope table](handbook.qmd#tbl-t3-slopes) that bound the curve on either side of the minimum. The marker is the optimum averaging time."
white_line <- adev[1] * sqrt(tau[1] / tau)
rw_line <- adev[length(adev)] * sqrt(tau / tau[length(tau)])
ggplot() +
  geom_line(data = data.frame(tau = tau, s = white_line),
            aes(tau, s), linetype = "dashed", linewidth = 0.4, colour = "grey50") +
  geom_line(data = data.frame(tau = tau, s = rw_line),
            aes(tau, s), linetype = "dashed", linewidth = 0.4, colour = "grey70") +
  geom_line(data = data.frame(tau = tau, s = adev), aes(tau, s), linewidth = 0.4) +
  geom_point(data = data.frame(tau = tau, s = adev), aes(tau, s), size = 1.6) +
  geom_vline(xintercept = tau_opt, linetype = "dotted", linewidth = 0.3, colour = "grey30") +
  annotate("point", x = tau_opt, y = sigma_opt, shape = 0, size = 4, stroke = 1.1) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = expression("Averaging time " * tau * " [s]"),
       y = expression(sigma[y](tau) * " [K]"))
```

```{r}
#| label: out-allan
#| echo: false
#| output: asis
rows <- sprintf("| %.1f | %.4f |", tau, adev * 1e3)
cat(sprintf('
| $\\tau$ [s] | $\\sigma_y(\\tau)$ [mK] |
|---|---|
%s

: Overlapping Allan deviation of the detrended residual {#tbl-t3-allan}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: out-allan-text
#| echo: false
#| output: asis
cat(sprintf('
The minimum is at $\\tau_\\mathrm{opt} = %.0f$ s
(%.1f min), where $\\sigma_y(\\tau_\\mathrm{opt}) =
%.4f$ mK. Below this the curve falls as white noise averages
down; above it, the random walk of [reading the slope](handbook.qmd#sec-t3-slope) takes over and averaging
longer makes the estimate of the mean *worse*.
', tau_opt, tau_opt / 60, sigma_opt * 1e3))
```

## $s/\sqrt{n}$ against the Allan deviation

The ordinary standard error of the mean, computed the way S1 computes it,
compared with the Allan deviation at the optimum.

```{r}
#| label: compare
s_naive <- sd(resid) / sqrt(n)
ratio <- s_naive / sigma_opt
```

```{r}
#| label: out-compare
#| echo: false
#| output: asis
cat(sprintf('
| Estimator | Value |
|---|---|
| $s/\\sqrt{n}$ over the whole record ($n = %d$) | %.4f mK |
| $\\sigma_y(\\tau_\\mathrm{opt})$ | %.4f mK |
| Ratio | %.2f |

: The naive estimator against the Allan-deviation optimum {#tbl-t3-compare}
', n, s_naive * 1e3, sigma_opt * 1e3, ratio))
```

```{r}
#| label: out-compare-text
#| echo: false
#| output: asis
verdict <- if (s_naive < sigma_opt) "understates" else "overstates"
cat(sprintf('
$s/\\sqrt n$ %s the true uncertainty of a long average of this record
by a factor of %.1f, because it assumes every one of the
$n$ samples is independent -- which [the Allan-deviation theory](handbook.qmd#sec-t3-allan) already showed is false at
long $\\tau$. $\\sigma_y(\\tau_\\mathrm{opt})$ is the honest number, and it
is the one the budget below uses.
', verdict, max(ratio, 1 / ratio)))
```

## Uncertainty budget

The measurand is [the mean temperature](handbook.qmd#eq-t3-that)
$\hat T = \bar x(\tau_\mathrm{opt}) - \Delta_\mathrm{cal}$, over the optimum
averaging window.

```{r}
#| label: effective-q
# S1's "effective quantization" idea, applied to this sensor chain: the
# Allan deviation at the shortest averaging time (tau0) is, by construction,
# an estimator of the single-sample (white) noise floor -- reusing @tbl-t3-allan
# rather than a fresh standard deviation, which would also pick up the
# random walk's contribution and overstate this term.
FSR_K <- (board$vref_mV / 1000) / channel$sensor_sensitivity_V_per_K
sigma_1 <- adev[1]
enob_eff <- enob(FSR_K, sigma_1)
q_eff_K <- FSR_K / 2^enob_eff
```

```{r}
#| label: budget
u_cal_K <- 0.05                    # TODO (Paolo): from T1, once run -- placeholder bound
drift_bound_K <- abs(drift_K_per_h / 3600) * tau_opt  # half-width, full drift over the window

components <- list(
  list(quantity = "Repeatability at the optimum", value = 0, unit = "K",
       distribution = "type-a", std = sigma_opt, sensitivity = 1.0),
  list(quantity = "Calibration (T1)", value = 0, unit = "K",
       distribution = "rectangular", half_width = u_cal_K, sensitivity = -1.0),
  list(quantity = "Effective quantization", value = 0, unit = "K",
       distribution = "rectangular", half_width = q_eff_K / 2, sensitivity = 1.0),
  list(quantity = "Drift over the window", value = 0, unit = "K",
       distribution = "rectangular", half_width = drift_bound_K / 2, sensitivity = 1.0)
)
budget <- budget_table(components, k = 2)
```

```{r}
#| label: out-budget
#| echo: false
#| output: asis
cat(budget_to_markdown(
  budget,
  caption = sprintf("Uncertainty budget for the mean temperature, $\\tau_\\mathrm{opt} = %.0f$ s", tau_opt),
  label = "tbl-t3-budget"))
```

```{r}
#| label: out-budget-text
#| echo: false
#| output: asis
dominant <- budget$table[which.max(budget$table$index_pct), ]
cat(sprintf('
Expanded uncertainty $U = %.3f$ mK at $k = 2$. The dominant
component is **%s** at %.0f %% of
the combined variance.
', budget$U * 1e3, dominant$quantity, dominant$index_pct))
```

::: {.callout-note}
## What this budget assumes

The calibration term uses a placeholder bound pending T1; replace
`u_cal_K` with T1's own reported uncertainty once that experiment has been
run. The drift term is a bound, not a correction, for the reason given in the
handbook: it would need to know where in the diurnal cycle the averaging
window sits, which is not assumed here.
:::

## What to report

1. @tbl-t3-fit and @fig-t3-fit: drift and diurnal amplitude, and which one
   dominates.
2. @fig-t3-allan and @tbl-t3-allan: the Allan-deviation curve, its slopes
   classified against [the slope table](handbook.qmd#tbl-t3-slopes), and the optimum averaging time.
3. @tbl-t3-compare: $s/\sqrt n$ against $\sigma_y(\tau_\mathrm{opt})$, and by
   how much they disagree.
4. @tbl-t3-budget: the budget, the dominant component, and what you would
   change first to improve it.
