---
title: "S2 — Analysis template (R)"
subtitle: "Dynamic characterization of the DAQ"
description: "Sine-fit ENOB, the timing penalty, aperture jitter and clock accuracy, in R"
engine: knitr
toc: true
toc-depth: 3
code-fold: false
---

This template runs end to end on the reference data shipped with the
experiment. To analyse your own measurements, replace the files in `data/`
and change the paths in the first chunk — nothing else.

The Python template [`analysis-py.qmd`](analysis-py.qmd) has the same
sections, the same figure labels and the same output tables, and produces the
same numbers from the same data.

## Setup

```{r}
#| label: setup
#| message: false
library(ggplot2)
library(purrr)
library(patchwork)

source("../../shared/R/labtools.R")

SWEEP_RUNS <- sort(Sys.glob("data/sweep-*.csv"))
DETAIL_RUN <- "data/sweep-0500hz.csv"   # the point examined in detail
MAX_FIT_ITERATIONS <- 50
```

```{r}
#| label: helpers
# The generator's setting for `quantity`, from the sidecar. The generator is
# an instrument used to make the measurement, not a channel on the board, so
# its settings live in references[].
generator <- function(meta, quantity) {
  for (ref in meta$references) {
    if (!is.null(ref$quantity) && ref$quantity == quantity) {
      return(list(value = ref$reference_value, u = ref$reference_u))
    }
  }
  stop(sprintf("no reference entry for '%s'", quantity))
}

# Four-parameter sine fit: A, B, C and the frequency. At a fixed frequency the
# model is linear in A, B and C, so each iteration solves the three-parameter
# problem and then takes one Gauss-Newton step in f using the extended design
# matrix. Written out rather than taken from a library so that R and Python
# follow the same arithmetic.
sine_fit4 <- function(t_s, y, f0) {
  n <- length(t_s)
  f <- f0
  for (i in seq_len(MAX_FIT_ITERATIONS)) {
    w <- 2 * pi * f * t_s
    d <- cbind(cos(w), sin(w), rep(1, n))
    coef <- solve(crossprod(d), crossprod(d, y))
    # d/df of the model, at the current parameters
    dcol <- 2 * pi * t_s * (-coef[1] * sin(w) + coef[2] * cos(w))
    e <- cbind(d, dcol)
    step <- solve(crossprod(e), crossprod(e, y))
    df <- as.numeric(step[4])
    if (!is.finite(df)) break
    f <- f + df
    if (abs(df) < 1e-12 * max(abs(f), 1)) break
  }

  w <- 2 * pi * f * t_s
  d <- cbind(cos(w), sin(w), rep(1, n))
  coef <- solve(crossprod(d), crossprod(d, y))
  resid <- as.vector(y - d %*% coef)
  dof <- max(n - 4, 1)
  sigma_r <- sqrt(sum(resid^2) / dof)

  # Standard error of the frequency, from the linearised 4-column problem.
  dcol <- 2 * pi * t_s * (-coef[1] * sin(w) + coef[2] * cos(w))
  e <- cbind(d, dcol)
  cov <- solve(crossprod(e))
  list(f = f,
       amplitude = sqrt(coef[1]^2 + coef[2]^2),
       level = as.numeric(coef[3]),
       sigma_r = sigma_r,
       u_f = sigma_r * sqrt(cov[4, 4]),
       residual = resid,
       coef = as.vector(coef))
}

# The nominal grid: what you would have if the samples were not stamped. The
# period comes from the least-squares grid fit, so this is the *best* uniform
# grid for the record -- the comparison is deliberately generous to the
# nominal-grid analysis.
uniform_time <- function(run) {
  t_us <- sort(unique(run$data$t_us))
  period_us <- timing_summary(run$data)$stats$period_us
  (t_us[1] + (seq_along(t_us) - 1) * period_us) / 1e6
}
```

```{r}
#| label: load
runs <- map(SWEEP_RUNS, read_run)
detail <- read_run(DETAIL_RUN)
board <- runs[[1]]$meta$board
FSR <- board$vref_mV / 1000
LSB <- lsb_volts(runs[[1]]$meta)
```

```{r}
#| label: banner
#| echo: false
#| output: asis
cat(synthetic_banner(runs[[1]]$meta))
```

## Data quality

```{r}
#| label: quality
#| echo: false
#| output: asis
issues <- map(runs, function(r) validate_meta(r$meta))
lossy <- character(0)
for (r in runs) {
  stats <- r$meta$stream_stats
  counters <- stats[setdiff(names(stats), "records")]
  if (any(map_lgl(counters, ~ is.numeric(.x) && .x > 0))) {
    lossy <- c(lossy, r$meta$run)
  }
}
n_bad <- sum(map_int(issues, length) > 0)
cat(sprintf('
%d sweep points read, %d samples each.
%s
%s
', length(runs), nrow(runs[[1]]$data),
   if (n_bad == 0) "All sidecars are complete."
   else sprintf("%d sidecar(s) have missing fields.", n_bad),
   if (length(lossy) == 0) "No run reports a lost record."
   else paste0("Runs with losses: ", paste(lossy, collapse = ", "))))
```

## One fit, in detail

Before sweeping anything, look at a single fit and satisfy yourself that the
model describes the data. A residual with structure in it means the number
you are about to compute is measuring your setup, not the converter.

```{r}
#| label: detail-fit
gen_f_detail <- generator(detail$meta, "frequency")
t_detail <- detail$data$t_us / 1e6
y_detail <- detail$data$volts

fit_detail <- sine_fit4(t_detail, y_detail, gen_f_detail$value)
```

```{r}
#| label: fig-s2-residual
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.0
#| fig-cap: "One sweep point. Top: the samples and the fitted sinusoid over the first few cycles. Bottom: the residual over the whole record. Structure here — flat tops, a slow trend, a visible harmonic — invalidates everything downstream."
t_rel_ms <- (t_detail - t_detail[1]) * 1e3
show <- t_rel_ms <= 3e3 / fit_detail$f          # about three cycles
dense <- seq(t_detail[show][1], tail(t_detail[show], 1), length.out = 800)
w <- 2 * pi * fit_detail$f * dense
model <- fit_detail$coef[1] * cos(w) + fit_detail$coef[2] * sin(w) +
  fit_detail$coef[3]

p_fit <- ggplot() +
  geom_line(data = data.frame(t = (dense - t_detail[1]) * 1e3, v = model),
            aes(t, v), linewidth = 0.4) +
  geom_point(data = data.frame(t = t_rel_ms[show], v = y_detail[show]),
             aes(t, v), size = 0.8) +
  labs(x = NULL, y = "ADC input (V)")

p_res <- ggplot(data.frame(t = t_rel_ms, r = fit_detail$residual * 1e3),
                aes(t, r)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.2) +
  labs(x = "Time (ms)", y = "Residual (mV)")

p_fit / p_res
```

```{r}
#| label: out-detail
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| Generator setting | %.4f Hz ± %.4f Hz |
| Fitted frequency $\\hat{f}$ | %.4f Hz |
| Standard error $u(\\hat{f})$ | %.3f mHz |
| Fitted amplitude | %.4f V |
| Fitted mean level | %.4f V |
| Residual $\\sigma_r$ | %.4f mV (%.1f LSB) |
| Effective bits | %.2f |
| Amplitude as a fraction of FSR | %.1f %% |

: The detailed fit, run `%s` {#tbl-s2-detail}
', gen_f_detail$value, gen_f_detail$u, fit_detail$f, fit_detail$u_f * 1e3,
   fit_detail$amplitude, fit_detail$level, fit_detail$sigma_r * 1e3,
   fit_detail$sigma_r / LSB, enob(FSR, fit_detail$sigma_r),
   2 * fit_detail$amplitude / FSR * 100, detail$meta$run))
```

## Effective bits in both time bases

The same samples, fitted twice. The only thing that changes is what is
assumed about *when* they were taken.

```{r}
#| label: sweep
sweep <- map(runs, function(run) {
  gf <- generator(run$meta, "frequency")
  ga <- generator(run$meta, "amplitude")
  y <- run$data$volts
  t_dev <- run$data$t_us / 1e6
  t_nom <- uniform_time(run)

  fit_dev <- sine_fit4(t_dev, y, gf$value)
  fit_nom <- sine_fit4(t_nom, y, gf$value)

  data.frame(
    run = run$meta$run,
    f_gen_Hz = gf$value, u_f_gen_Hz = gf$u,
    a_gen_V = ga$value, u_a_gen_V = ga$u,
    f_hat_Hz = fit_dev$f, u_f_hat_Hz = fit_dev$u_f,
    amplitude_V = fit_dev$amplitude,
    sigma_dev_V = fit_dev$sigma_r,
    sigma_nom_V = fit_nom$sigma_r,
    enob_dev = enob(FSR, fit_dev$sigma_r),
    enob_nom = enob(FSR, fit_nom$sigma_r),
    ppm = (fit_dev$f / gf$value - 1) * 1e6,
    n = length(y),
    duration_s = t_dev[length(t_dev)] - t_dev[1])
}) |> list_rbind()
sweep <- sweep[order(sweep$f_gen_Hz), ]
rownames(sweep) <- NULL
sweep$penalty_bits <- sweep$enob_dev - sweep$enob_nom
```

```{r}
#| label: out-sweep
#| echo: false
#| output: asis
rows <- sprintf("| %.2f | %.4f | %.4f | %.2f | %.4f | %.2f | %.2f |",
                sweep$f_gen_Hz, sweep$amplitude_V, sweep$sigma_dev_V * 1e3,
                sweep$enob_dev, sweep$sigma_nom_V * 1e3, sweep$enob_nom,
                sweep$penalty_bits)
cat(sprintf('
| $f_\\mathrm{gen}$ (Hz) | Amplitude (V) | $\\sigma_r$ on `t_us` (mV) | ENOB | $\\sigma_r$ on the grid (mV) | ENOB | Penalty (bit) |
|---|---|---|---|---|---|---|
%s

: Effective bits from the device timestamps and from the nominal grid {#tbl-s2-sweep}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: out-sweep-text
#| echo: false
#| output: asis
lo <- sweep[1, ]
hi <- sweep[nrow(sweep), ]
cat(sprintf('
At %.0f Hz the converter delivers %.2f bits;
at %.0f Hz, %.2f. The loss of
%.2f bits across the sweep is the timing of
the sample-and-hold, not its noise — the noise has not changed.

The last column is the result this experiment exists for. Analysing the
**same samples** on the nominal grid costs between
%.1f and %.1f bits.
That is what the device\'s timestamps are worth, and what an acquisition
system that does not report them silently takes away.
', lo$f_gen_Hz, lo$enob_dev, hi$f_gen_Hz, hi$enob_dev,
   lo$enob_dev - hi$enob_dev, min(sweep$penalty_bits), max(sweep$penalty_bits)))
```

## The roll-off and the jitter limit

Two mechanisms add in quadrature: a noise floor that does not depend on
frequency, and a jitter term that grows with it.

$$
\sigma_r^2(f) = \sigma_n^2
  + \left(\frac{2\pi f A \sigma_j}{\sqrt{2}}\right)^2
$$ {#eq-s2-quadrature}

which is linear in $f^2$, so both $\sigma_n$ and $\sigma_j$ come from one
straight-line fit rather than from eyeballing the high-frequency end.

The $\sqrt{2}$ is not decoration. A timing error $\delta t$ produces an
amplitude error $\dot{V}\delta t$, and the slope of a sinusoid is
$\dot{V} = 2\pi f A\cos\theta$, whose RMS over the cycle is $2\pi f A/\sqrt{2}$
— not its peak. Dropping it reports an aperture jitter $\sqrt{2}$ times too
small, and it is the same $\sqrt{2}$ that cancels in the standard
$\mathrm{SNR}_j = -20\log_{10}(2\pi f\sigma_j)$, where signal and error are
both expressed as RMS.

```{r}
#| label: jitter
x <- sweep$f_gen_Hz^2
y <- sweep$sigma_dev_V^2
sxx <- sum((x - mean(x))^2)
slope <- sum((x - mean(x)) * (y - mean(y))) / sxx
intercept <- mean(y) - slope * mean(x)

amp_mean <- mean(sweep$amplitude_V)
sigma_n <- sqrt(max(intercept, 0))
# slope = (2 pi A sigma_j / sqrt(2))^2, so the sqrt(2) comes back here.
sigma_j <- sqrt(2) * sqrt(max(slope, 0)) / (2 * pi * amp_mean)

# Where the two mechanisms are equal.
f_cross <- sqrt(2) * sigma_n / (2 * pi * amp_mean * sigma_j)
enob_noise <- enob(FSR, sigma_n)
```

```{r}
#| label: fig-s2-enob
#| echo: false
#| fig-width: 7.0
#| fig-height: 4.2
#| fig-cap: "Effective bits against frequency. Circles: from the device timestamps. Crosses: the same samples on the nominal grid. Solid: the two-mechanism model of @eq-s2-quadrature. Dashed: the noise floor alone and the jitter limit alone; they cross where the converter stops being noise limited."
f_grid <- 10^seq(log10(min(sweep$f_gen_Hz) * 0.7),
                 log10(max(sweep$f_gen_Hz) * 1.5), length.out = 300)
model_sigma <- sqrt(sigma_n^2 + (2 * pi * f_grid * amp_mean * sigma_j / sqrt(2))^2)
# enob() is scalar in both languages, so it is mapped over the grid here
# exactly as the Python template maps it in a comprehension.
enob_v <- function(sigmas) map_dbl(sigmas, ~ enob(FSR, .x))

curves <- rbind(
  data.frame(f = f_grid, bits = enob_v(model_sigma),
             what = "noise and jitter combined"),
  data.frame(f = f_grid, bits = rep(enob_noise, length(f_grid)),
             what = sprintf("noise floor alone (%.2f bit)", enob_noise)),
  data.frame(f = f_grid,
             bits = enob_v(2 * pi * f_grid * amp_mean * sigma_j / sqrt(2)),
             what = "jitter limit alone"))
points <- rbind(
  data.frame(f = sweep$f_gen_Hz, bits = sweep$enob_dev, what = "from t_us"),
  data.frame(f = sweep$f_gen_Hz, bits = sweep$enob_nom,
             what = "on the nominal grid"))

ggplot() +
  geom_vline(xintercept = f_cross, linewidth = 0.3, linetype = "dotted",
             colour = "grey40") +
  geom_line(data = curves, aes(f, bits, linetype = what), linewidth = 0.4) +
  geom_point(data = points, aes(f, bits, shape = what), size = 2) +
  scale_x_log10() +
  coord_cartesian(ylim = c(0, enob_noise + 2)) +
  labs(x = "Generator frequency (Hz)", y = "Effective number of bits",
       linetype = NULL, shape = NULL)
```

```{r}
#| label: out-jitter
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| Noise floor $\\sigma_n$ | %.4f mV (%.1f LSB) |
| Noise-limited ENOB | %.2f bit |
| Aperture jitter $\\sigma_j$ | %.1f ns |
| Mean fitted amplitude | %.4f V |
| Crossover frequency | %.0f Hz |
| ENOB at the crossover | %.2f bit |

: Separating the two mechanisms, from the straight-line fit of @eq-s2-quadrature {#tbl-s2-jitter}
', sigma_n * 1e3, sigma_n / LSB, enob_noise, sigma_j * 1e9, amp_mean,
   f_cross, enob(FSR, sqrt(2) * sigma_n)))
```

```{r}
#| label: out-jitter-text
#| echo: false
#| output: asis
cat(sprintf('
Below **%.0f Hz** the converter is noise limited: a quieter front end
would improve it and a better clock would not. Above it the converter is
jitter limited, and no amount of front-end work helps — @eq-s2-quadrature\'s
second term does not contain the number of bits, so a 16-bit converter and a
12-bit one with the same %.0f ns of aperture jitter reach the
same ceiling.

Compare $\\sigma_n = %.3f$ mV with the noise floor measured on a
constant input in [S1](../S1-daq-static/index.qmd). They are the same quantity
measured two different ways, and they should agree. The low-frequency dynamic
ENOB, %.2f bit, can never exceed S1\'s static figure —
if it does, one of the two experiments is wrong.
', f_cross, sigma_j * 1e9, sigma_n * 1e3, max(sweep$enob_dev)))
```

## Clock accuracy

The fitted frequency is expressed in the device's time base; the generator's
setting is expressed in the generator's. Their ratio compares two oscillators.

```{r}
#| label: clock
sweep$u_ppm_fit <- sweep$u_f_hat_Hz / sweep$f_gen_Hz * 1e6
sweep$u_ppm_gen <- sweep$u_f_gen_Hz / sweep$f_gen_Hz * 1e6
ppm_mean <- mean(sweep$ppm)
ppm_spread <- max(sweep$ppm) - min(sweep$ppm)
```

```{r}
#| label: out-clock
#| echo: false
#| output: asis
rows <- sprintf("| %.2f | %.4f | %+.2f | %.3f | %.2f |",
                sweep$f_gen_Hz, sweep$f_hat_Hz, sweep$ppm,
                sweep$u_ppm_fit, sweep$u_ppm_gen)
cat(sprintf('
| $f_\\mathrm{gen}$ (Hz) | $\\hat{f}$ (Hz) | $\\delta$ (ppm) | $u(\\hat{f})$ (ppm) | $u(f_\\mathrm{gen})$ (ppm) |
|---|---|---|---|---|
%s

: Clock accuracy at every sweep point {#tbl-s2-clock}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: fig-s2-ppm
#| echo: false
#| fig-width: 7.0
#| fig-height: 3.2
#| fig-cap: "Clock error against frequency, with the standard error of each fit. The quantity should not depend on frequency: a trend would mean the fit is not converging, not that the oscillator changes with the signal."
band <- mean(sweep$u_ppm_gen)
ggplot(sweep, aes(f_gen_Hz, ppm)) +
  annotate("rect", xmin = min(sweep$f_gen_Hz) * 0.7,
           xmax = max(sweep$f_gen_Hz) * 1.5, ymin = -band, ymax = band,
           fill = "grey85") +
  geom_hline(yintercept = ppm_mean, linewidth = 0.3, linetype = "dashed",
             colour = "grey40") +
  geom_errorbar(aes(ymin = ppm - u_ppm_fit, ymax = ppm + u_ppm_fit),
                width = 0.04, linewidth = 0.3) +
  geom_point(size = 1.8) +
  scale_x_log10() +
  labs(x = "Generator frequency (Hz)", y = "Clock error (ppm)",
       subtitle = sprintf("mean %+.2f ppm; grey band = generator specification",
                          ppm_mean))
```

```{r}
#| label: out-clock-text
#| echo: false
#| output: asis
sign_word <- if (ppm_mean < 0) "fast" else "slow"
resolved_gen <- if (abs(ppm_mean) > mean(sweep$u_ppm_gen)) "larger than" else "comparable to"
cat(sprintf('
Mean clock error **%+.2f ppm**, with a spread of
%.2f ppm across two and a half decades of signal frequency and no
trend — which is what it should be, since an oscillator\'s rate does not depend
on what you measure with it.

The sign says the device\'s oscillator runs **%s**: a device clock running
fast stamps a given real interval with too many microseconds, so a real
frequency appears low when expressed in device time.

Against a generator specified to
±%.1f ppm, this result is
%s
the reference\'s own uncertainty — see the budget below before concluding
anything about the board.
', ppm_mean, ppm_spread, sign_word, mean(sweep$u_ppm_gen), resolved_gen))
```

## Uncertainty budget

The measurand is the clock error in ppm, with the
[sensitivity coefficients](handbook.qmd#eq-s2-ppm-sensitivity) of the
handbook. It is evaluated at the sweep point with the smallest fit
uncertainty.

```{r}
#| label: budget
best <- sweep[which.min(sweep$u_ppm_fit), ]
c_ppm <- 1e6 / best$f_gen_Hz

# One microsecond of timestamp quantization over the whole record, as a bound.
u_q_ppm <- 1e6 / (best$duration_s * 1e6)

components <- list(
  list(quantity = "Fitted frequency", value = best$f_hat_Hz, unit = "ppm",
       distribution = "type-a", std = best$u_f_hat_Hz, sensitivity = c_ppm),
  list(quantity = "Generator frequency", value = best$f_gen_Hz, unit = "ppm",
       distribution = "type-a", std = best$u_f_gen_Hz, sensitivity = -c_ppm),
  list(quantity = "Timestamp quantization", value = 0, unit = "ppm",
       distribution = "rectangular",
       half_width = u_q_ppm * best$f_gen_Hz / 1e6, sensitivity = c_ppm),
  list(quantity = "Instability across the sweep", value = 0, unit = "ppm",
       distribution = "rectangular",
       half_width = ppm_spread / 2 * best$f_gen_Hz / 1e6, sensitivity = c_ppm)
)
budget <- budget_table(components, k = 2)
```

```{r}
#| label: out-budget
#| echo: false
#| output: asis
cat(budget_to_markdown(
  budget,
  caption = sprintf("Uncertainty budget for the clock error, evaluated at $f_\\mathrm{gen} = %.2f$ Hz", best$f_gen_Hz),
  label = "tbl-s2-budget"))
```

```{r}
#| label: out-budget-text
#| echo: false
#| output: asis
dominant <- budget$table[which.max(budget$table$index_pct), ]
resolved <- abs(best$ppm) > budget$U
verdict <- if (resolved) {
  "The result is resolved: the measured offset is larger than its own expanded uncertainty, so the board s oscillator really does differ from the generator s by about this much."
} else {
  sprintf("The result is NOT resolved: the measured offset is smaller than its own expanded uncertainty, so all that can honestly be said is that the two oscillators agree to within %.1f ppm.", budget$U)
}
cat(sprintf('
Clock error $\\delta = %+.2f$ ppm with an expanded uncertainty
$U = %.2f$ ppm at $k = 2$. The dominant component is
**%s** at %.0f %% of the combined
variance.

%s

Note which component dominates. Improving this measurement means a better
*generator* or a longer record — not a better board. Being able to say which
is what the budget is for.
', best$ppm, budget$U, dominant$quantity, dominant$index_pct, verdict))
```

## What to report

1. @tbl-s2-detail and @fig-s2-residual: one fit examined closely, with a
   residual you are prepared to defend as structureless.
2. @tbl-s2-sweep: effective bits in both time bases, and the timing penalty.
   State the penalty as a sentence about acquisition systems in general.
3. @fig-s2-enob and @tbl-s2-jitter: the roll-off, the extracted $\sigma_n$ and
   $\sigma_j$, and the crossover. Compare $\sigma_n$ with S1's noise floor and
   say whether they agree.
4. @tbl-s2-clock and @fig-s2-ppm: the ppm result, its sign explained, and
   whether it depends on frequency.
5. @tbl-s2-budget: the budget, the dominant component, and what you would
   change to improve the measurement.
