---
title: "S3 — Analysis template (R)"
subtitle: "Aliasing, conditioning and filtering"
description: "Folding predicted and observed, filter attenuation, and what digital filtering cannot undo, in R"
engine: knitr
toc: true
toc-depth: 3
code-fold: false
---

This template runs end to end on the reference data shipped with the
experiment. To analyse your own measurements, replace the files in `data/`
and change the paths in the first chunk — nothing else.

The Python template [`analysis-py.qmd`](analysis-py.qmd) has the same
sections, the same figure labels and the same output tables, and produces the
same numbers from the same data.

::: {.callout-note collapse="true"}
## Why the filters here are written out rather than called from a library

Both templates apply the same
[windowed-sinc FIR](handbook.qmd#eq-s3-fir), computed from its closed form and
applied by direct convolution. Library filter design and `filtfilt` edge
padding differ between `scipy.signal` and `gsignal` by enough to break
table-for-table agreement between the two languages, and an analysis whose
numbers depend on which language ran it is not an analysis. The frequency
estimator and the sine fit are written out for the same reason.
:::

## Setup

```{r}
#| label: setup
#| message: false
library(ggplot2)
library(patchwork)

source("../../shared/R/labtools.R")

theme_set(theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.25)))

SWEEP_RUNS <- sort(Sys.glob("data/alias-f*.csv"))
FILTER_OFF <- "data/filter-off.csv"
FILTER_ON <- "data/filter-on.csv"
OVERSAMPLED <- "data/oversampled.csv"

FIR_TAPS <- 201           # odd, so the group delay is an integer sample;
                          # enough taps that 150 Hz is inside the passband
                          # rather than in the transition band
DECIMATION <- 10          # 5 kHz oversampled run down to the 500 Hz grid
```

```{r}
#| label: helpers
# Hz, from the least-squares grid fit of the device timestamps. Not the
# requested rate and not achieved_rate_hz: the fitted period is the one both
# languages compute identically, and it is what the signal actually saw.
sampling_rate <- function(run) 1e6 / timing_summary(run$data)$stats$period_us

# sin(pi x)/(pi x), and 1 at x = 0 -- written out because R has no sinc().
sinc <- function(x) ifelse(x == 0, 1, sin(pi * x) / (pi * x))

# Peak of the Hann-windowed periodogram, refined by parabolic interpolation
# on the three bins around the peak.
estimate_frequency <- function(x, fs) {
  n <- length(x)
  w <- 0.5 - 0.5 * cos(2 * pi * (0:(n - 1)) / (n - 1))
  spec <- fft((x - mean(x)) * w)[1:(n %/% 2 + 1)]
  mag <- Mod(spec)
  k <- which.max(mag[-1]) + 1L          # 1-based; skip DC
  delta <- 0
  if (k > 1 && k < length(mag)) {
    a <- mag[k - 1]; b <- mag[k]; cc <- mag[k + 1]
    denom <- a - 2 * b + cc
    if (denom != 0) delta <- 0.5 * (a - cc) / denom
  }
  ((k - 1) + delta) * fs / n            # k - 1: back to the 0-based bin index
}

# Three-parameter least squares at a known frequency. Returns the amplitude,
# the mean level and the residual standard deviation. The normal equations are
# solved explicitly so that R and Python take the same arithmetic path.
sine_fit <- function(t_s, x, f) {
  d <- cbind(cos(2 * pi * f * t_s), sin(2 * pi * f * t_s), rep(1, length(t_s)))
  coef <- solve(crossprod(d), crossprod(d, x))
  resid <- as.vector(x - d %*% coef)
  dof <- max(length(t_s) - 3, 1)
  list(amplitude = sqrt(coef[1]^2 + coef[2]^2),
       level = as.numeric(coef[3]),
       resid_sd = sqrt(sum(resid^2) / dof))
}

# The folding rule: f_a = |f - k fs| with k the nearest integer. R's round()
# and Python's round() are both round-half-to-even, so the two agree; the
# sweep avoids half-integer f/fs anyway.
alias_frequency <- function(f, fs) abs(f - round(f / fs) * fs)

# Windowed-sinc low-pass, normalised to unity passband gain.
fir_lowpass <- function(fc, fs, n_taps = FIR_TAPS) {
  n <- 0:(n_taps - 1)
  m <- (n_taps - 1) / 2
  w <- 0.54 - 0.46 * cos(2 * pi * n / (n_taps - 1))    # Hamming
  h <- w * sinc(2 * fc / fs * (n - m))
  h / sum(h)
}

# Direct convolution, kept only where the whole window exists -- the same
# samples numpy's convolve(..., mode = "valid") returns.
fir_apply <- function(x, h) {
  y <- stats::filter(x, h, method = "convolution", sides = 1)
  as.vector(y[length(h):length(x)])
}

# |H(f)| of the Butterworth-style stage.
filter_magnitude <- function(f, fc, order) 1 / sqrt(1 + (f / fc)^(2 * order))
```

```{r}
#| label: load
sweep <- lapply(SWEEP_RUNS, read_run)
off <- read_run(FILTER_OFF)
on <- read_run(FILTER_ON)
over <- read_run(OVERSAMPLED)

all_runs <- c(sweep, list(off, on, over))
names(all_runs) <- vapply(all_runs, function(r) r$meta$run, character(1))
board <- on$meta$board
FSR <- board$vref_mV / 1000
LSB <- lsb_volts(on$meta)
```

```{r}
#| label: banner
#| echo: false
#| output: asis
cat(synthetic_banner(on$meta))
```

## Data quality

```{r}
#| label: quality
#| echo: false
#| output: asis
problems <- lapply(all_runs, function(r) validate_meta(r$meta))
lossy <- character(0)
for (name in names(all_runs)) {
  stats <- all_runs[[name]]$meta$stream_stats
  counters <- stats[setdiff(names(stats), "records")]
  if (any(vapply(counters, function(v) is.numeric(v) && v > 0, logical(1)))) {
    lossy <- c(lossy, name)
  }
}
n_bad <- sum(vapply(problems, length, integer(1)) > 0)
cat(sprintf('
%d runs read: %d in the aliasing sweep, the
filtered/unfiltered pair, and the oversampled run.
%s
%s
', length(all_runs), length(sweep),
   if (n_bad == 0) "All sidecars are complete."
   else sprintf("%d sidecar(s) have missing fields.", n_bad),
   if (length(lossy) == 0) "No run reports a lost record."
   else paste0("Runs with losses: ", paste(lossy, collapse = ", "))))
```

## The conditioning stage

The stage maps the generator's bipolar output onto the converter's unipolar
input; see the
[derivation in the handbook](handbook.qmd#sec-s3-conditioning). Its design
values live in each sidecar's `conditioning` block, and the check below is
that the recordings sit where the design says they should.

```{r}
#| label: conditioning
ch_on <- on$meta$channels[[1]]
GAIN <- ch_on$conditioning$gain
OFFSET <- ch_on$conditioning$offset
U_GAIN_REL <- ch_on$conditioning_u_gain_rel
U_OFFSET_V <- ch_on$conditioning_u_offset_v

cond <- do.call(rbind, lapply(list(off, on, over), function(run) {
  data.frame(run = run$meta$run,
             mean_V = mean(run$data$volts),
             min_V = min(run$data$volts),
             max_V = max(run$data$volts))
}))
cond$offset_error_mV <- (cond$mean_V - OFFSET) * 1e3
cond$headroom_lo_V <- cond$min_V
cond$headroom_hi_V <- FSR - cond$max_V
```

```{r}
#| label: out-conditioning
#| echo: false
#| output: asis
rows <- sprintf("| %s | %.4f | %+.1f | %.4f | %.4f | %.3f | %.3f |",
                cond$run, cond$mean_V, cond$offset_error_mV, cond$min_V,
                cond$max_V, cond$headroom_lo_V, cond$headroom_hi_V)
cat(sprintf('
| Run | Mean [V] | Mean − $V_\\mathrm{off}$ [mV] | Min [V] | Max [V] | Low headroom [V] | High headroom [V] |
|---|---|---|---|---|---|---|
%s

: Where each recording sits in the converter\'s range, against a design offset of %.3f V and a gain of %.3f {#tbl-s3-conditioning}
', paste(rows, collapse = "\n"), OFFSET, GAIN))
```

```{r}
#| label: out-conditioning-text
#| echo: false
#| output: asis
worst <- cond[which.max(abs(cond$offset_error_mV)), ]
cat(sprintf('
No recording reaches either rail: the smallest low headroom is
%.3f V and the smallest high headroom is
%.3f V, so nothing clips and the sinusoid fits
are unbiased. The largest departure of a recording\'s mean from the design
offset is %+.1f mV on `%s`, which is the
stage\'s own offset error plus the converter\'s — the two are not separable from
this measurement alone, which is why the budget carries an offset term from
the sidecar rather than from this table.
', min(cond$headroom_lo_V), min(cond$headroom_hi_V),
   worst$offset_error_mV, worst$run))
```

## Aliasing sweep

Each run is one generator frequency, sampled deliberately below Nyquist. The
apparent frequency is estimated from the recording and predicted from the
[folding rule](handbook.qmd#eq-s3-alias) using the generator setting recorded
in the sidecar.

```{r}
#| label: aliasing
alias <- do.call(rbind, lapply(sweep, function(run) {
  ch <- run$meta$channels[[1]]
  fs <- sampling_rate(run)
  volts <- run$data$volts
  t_s <- run$data$t_us / 1e6
  f_gen <- ch$generator_frequency_hz
  f_obs <- estimate_frequency(volts, fs)
  f_pred <- alias_frequency(f_gen, fs)
  fit <- sine_fit(t_s, volts, f_obs)
  data.frame(run = run$meta$run, f_gen_Hz = f_gen, fs_Hz = fs,
             k = round(f_gen / fs), f_pred_Hz = f_pred, f_obs_Hz = f_obs,
             error_Hz = f_obs - f_pred, amplitude_V = fit$amplitude)
}))
```

```{r}
#| label: out-alias
#| echo: false
#| output: asis
rows <- sprintf("| %.0f | %d | %.2f | %.2f | %+.3f | %.4f |",
                alias$f_gen_Hz, alias$k, alias$f_pred_Hz, alias$f_obs_Hz,
                alias$error_Hz, alias$amplitude_V)
cat(sprintf('
| $f$ [Hz] | $k$ | Predicted $f_a$ [Hz] | Observed [Hz] | Error [Hz] | Amplitude [V] |
|---|---|---|---|---|---|
%s

: The aliasing sweep at $f_s = %.1f$ Hz {#tbl-s3-alias}
', paste(rows, collapse = "\n"), alias$fs_Hz[1]))
```

```{r}
#| label: fig-s3-folding
#| echo: false
#| fig-width: 7.6
#| fig-height: 3.4
#| fig-cap: "The folding triangle. Left: apparent frequency against generator frequency, predicted (line) and observed (points); every point above the Nyquist frequency is a different generator setting producing the same samples as something below it. Right: observed minus predicted."
fs0 <- alias$fs_Hz[1]
grid <- seq(0, max(alias$f_gen_Hz) * 1.02, length.out = 2000)
pred <- data.frame(f = grid, fa = alias_frequency(grid, fs0))

p_fold <- ggplot() +
  geom_line(data = pred, aes(f, fa), linewidth = 0.4) +
  geom_point(data = alias, aes(f_gen_Hz, f_obs_Hz), size = 1.8) +
  geom_hline(yintercept = fs0 / 2, linewidth = 0.3, linetype = "dashed",
             colour = "grey40") +
  annotate("text", x = max(grid), y = fs0 / 2, label = " Nyquist",
           hjust = 1, vjust = -0.6, size = 2.8, colour = "grey30") +
  labs(x = "Generator frequency $f$ [Hz]", y = "Apparent frequency [Hz]")

p_err <- ggplot(alias, aes(f_gen_Hz, error_Hz * 1e3)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.3) +
  geom_point(size = 1.6) +
  labs(x = "Generator frequency [Hz]", y = "Observed − predicted [mHz]")

p_fold | p_err
```

```{r}
#| label: out-alias-text
#| echo: false
#| output: asis
cat(sprintf('
The largest departure from the folding rule across the sweep is
%.1f mHz, against a bin width of
%.2f Hz. The rule is exact; what
is left is the frequency estimator\'s own resolution.

Read the left panel again and note what it means at the bench: every one of
the %d points with $k > 0$ is a generator setting above
Nyquist that produced samples identical to those of some lower frequency.
Nothing in the recorded numbers distinguishes them.
', max(abs(alias$error_Hz)) * 1e3,
   alias$fs_Hz[1] / nrow(sweep[[1]]$data), sum(alias$k > 0)))
```

## Anti-alias filtering

The same 650 Hz input, recorded twice: once straight into the converter, once
through the analog filter. Both alias to the same apparent frequency — the
filter does not move the alias, it reduces what arrives to be aliased.

```{r}
#| label: filtering
FC <- ch_on$filter_fc_hz
ORDER <- ch_on$filter_order
U_FC_REL <- ch_on$filter_u_fc_rel
F_GEN <- ch_on$generator_frequency_hz

pair <- do.call(rbind, Map(function(label, run) {
  fs <- sampling_rate(run)
  volts <- run$data$volts
  t_s <- run$data$t_us / 1e6
  f_obs <- estimate_frequency(volts, fs)
  fit <- sine_fit(t_s, volts, f_obs)
  data.frame(case = label, fs_Hz = fs, f_obs_Hz = f_obs,
             amplitude_V = fit$amplitude, level_V = fit$level,
             resid_V = fit$resid_sd, n = length(volts))
}, c("unfiltered", "filtered"), list(off, on)))

ratio_obs <- pair$amplitude_V[pair$case == "filtered"] /
  pair$amplitude_V[pair$case == "unfiltered"]
ratio_pred <- filter_magnitude(F_GEN, FC, ORDER)
```

```{r}
#| label: out-filter
#| echo: false
#| output: asis
rows <- sprintf("| %s | %.2f | %.5f | %.4f | %.3f | %d |",
                pair$case, pair$f_obs_Hz, pair$amplitude_V, pair$level_V,
                pair$resid_V * 1e3, pair$n)
cat(sprintf('
| Case | Observed $f_a$ [Hz] | Amplitude [V] | Mean level [V] | Residual $\\sigma$ [mV] | $n$ |
|---|---|---|---|---|---|
%s

: The same %.0f Hz input with and without the analog filter {#tbl-s3-filter}
', paste(rows, collapse = "\n"), F_GEN))
```

```{r}
#| label: out-filter-text
#| echo: false
#| output: asis
cat(sprintf('
Observed amplitude ratio %.4f
(%+.2f dB); predicted from
$\\lvert H(f)\\rvert$ at the **true** %.0f Hz — not at the apparent
%.0f Hz — with $f_c = %.0f$ Hz and
$n = %d$: %.4f (%+.2f dB).
The difference is %+.2f %%.

Both recordings still show a tone at %.0f Hz. That
is the point: the filter changed the alias\'s *size*, not its existence or its
frequency. Attenuating it enough is a design decision taken before the
sampler, and the next section is how that decision is made.
', ratio_obs, 20 * log10(ratio_obs), F_GEN, pair$f_obs_Hz[1], FC, ORDER,
   ratio_pred, 20 * log10(ratio_pred), 100 * (ratio_obs / ratio_pred - 1),
   pair$f_obs_Hz[2]))
```

```{r}
#| label: fig-s3-filter-pair
#| echo: false
#| fig-width: 7.0
#| fig-height: 3.2
#| fig-cap: "First 60 ms of the unfiltered and filtered recordings. The apparent frequency is identical — the filter cannot move an alias — and only the amplitude differs."
trace <- do.call(rbind, Map(function(label, run) {
  t_ms <- (run$data$t_us - run$data$t_us[1]) / 1e3
  keep <- t_ms <= 60
  data.frame(case = label, t_ms = t_ms[keep], volts = run$data$volts[keep])
}, c("unfiltered", "filtered"), list(off, on)))

ggplot(trace, aes(t_ms, volts, colour = case)) +
  geom_hline(yintercept = OFFSET, linewidth = 0.3, linetype = "dashed",
             colour = "grey50") +
  geom_line(linewidth = 0.3) +
  geom_point(size = 0.6) +
  labs(x = "Time [ms]", y = "ADC input [V]", colour = NULL) +
  theme(legend.position = "top")
```

## Filter design

The design question of
[the handbook](handbook.qmd#sec-s3-filter-design): given a sampling rate, a
passband edge and a required attenuation at the worst-case folding frequency,
what corner frequency and what order? The cost is paid in the passband.

```{r}
#| label: design
FS_DESIGN <- 500       # the rate the bench actually samples at
F_MAX <- 100           # the highest frequency of interest
A_REQ_DB <- 40         # required attenuation at the folding frequency
F_FOLD <- FS_DESIGN - F_MAX

design <- do.call(rbind, lapply(1:4, function(order) {
  fc <- F_FOLD / (10^(A_REQ_DB / 10) - 1)^(1 / (2 * order))
  droop <- filter_magnitude(F_MAX, fc, order)
  data.frame(order = order, fc_Hz = fc, fc_over_fmax = fc / F_MAX,
             droop_dB = 20 * log10(droop), droop_pct = 100 * (1 - droop))
}))
```

```{r}
#| label: out-design
#| echo: false
#| output: asis
rows <- sprintf("| %d | %.1f | %.2f | %+.2f | %.1f |",
                design$order, design$fc_Hz, design$fc_over_fmax,
                design$droop_dB, design$droop_pct)
cat(sprintf('
| Order $n$ | Required $f_c$ [Hz] | $f_c/f_\\mathrm{max}$ | Droop at $f_\\mathrm{max}$ [dB] | Droop [%%] |
|---|---|---|---|---|
%s

: Corner frequency needed for %.0f dB at $f_\\mathrm{fold} = %.0f$ Hz, with $f_s = %.0f$ Hz and $f_\\mathrm{max} = %.0f$ Hz {#tbl-s3-design}
', paste(rows, collapse = "\n"), A_REQ_DB, F_FOLD, FS_DESIGN, F_MAX))
```

```{r}
#| label: fig-s3-design
#| echo: false
#| fig-width: 7.0
#| fig-height: 3.6
#| fig-cap: "Magnitude response of each design of @tbl-s3-design. Every curve passes through the required attenuation at the folding frequency; they differ in how much of the passband they spend to get there."
f_grid <- 10^seq(0, log10(FS_DESIGN * 2), length.out = 400)
curves <- do.call(rbind, lapply(seq_len(nrow(design)), function(i) {
  data.frame(design_label = sprintf("n = %d, fc = %.0f Hz", design$order[i],
                                    design$fc_Hz[i]),
             f = f_grid,
             mag_dB = 20 * log10(filter_magnitude(f_grid, design$fc_Hz[i],
                                                  design$order[i])))
}))

ggplot(curves, aes(f, mag_dB, colour = design_label)) +
  geom_vline(xintercept = F_MAX, linewidth = 0.3, linetype = "dotted",
             colour = "grey40") +
  geom_vline(xintercept = F_FOLD, linewidth = 0.3, linetype = "dashed",
             colour = "grey40") +
  geom_hline(yintercept = -A_REQ_DB, linewidth = 0.3, linetype = "dashed",
             colour = "grey40") +
  geom_line(linewidth = 0.4) +
  scale_x_log10() +
  coord_cartesian(ylim = c(-70, 12)) +
  labs(x = "Frequency [Hz]", y = "|H(f)| [dB]", colour = NULL) +
  theme(legend.position = "bottom", legend.text = element_text(size = 7))
```

```{r}
#| label: out-design-text
#| echo: false
#| output: asis
first <- design[1, ]
last <- design[nrow(design), ]
cat(sprintf('
A first-order stage needs $f_c = %.1f$ Hz to reach
%.0f dB at %.0f Hz — below the passband edge itself, which
costs %.0f %% of the signal at $f_\\mathrm{max}$ and is
not a usable design. Order %.0f puts the corner at
%.1f Hz, %.1f times the passband edge,
for %.1f %% droop. That is the trade the table exists to
make visible: attenuation is bought with either order or passband.
', first$fc_Hz, A_REQ_DB, F_FOLD, first$droop_pct, last$order, last$fc_Hz,
   last$fc_over_fmax, last$droop_pct))
```

## Digital filtering after the fact

The oversampled run contains the signal and a strong interferer well above
it, at a rate high enough that neither folds. Filtering *then* decimating is
clean. Decimating without filtering first is not — and no filter applied
afterwards repairs it.

```{r}
#| label: digital
ch_over <- over$meta$channels[[1]]
F_SIGNAL <- ch_over$signal_frequency_hz
F_INTERFERER <- ch_over$interferer_frequency_hz

fs_over <- sampling_rate(over)
v_over <- over$data$volts
t_over <- over$data$t_us / 1e6
fs_dec <- fs_over / DECIMATION

# The digital anti-alias filter: corner below the interferer, above the signal.
FC_DIGITAL <- fs_dec / 2 * 0.8
h <- fir_lowpass(FC_DIGITAL, fs_over)
lag <- (length(h) - 1) %/% 2

v_filt <- fir_apply(v_over, h)
t_filt <- t_over[(lag + 1):(length(v_over) - lag)]

# Where the interferer lands if it is allowed to fold onto the decimated grid.
F_FOLDED <- alias_frequency(F_INTERFERER, fs_dec)

# 1. naive: decimate the oversampled run with no digital filter
idx1 <- seq(1, length(v_over), by = DECIMATION)
# 2. filter first, then decimate
idx2 <- seq(1, length(v_filt), by = DECIMATION)
# 3. the already-aliased run, filtered digitally
fs_off <- sampling_rate(off)
h_off <- fir_lowpass(fs_off / 2 * 0.8, fs_off)
lag_off <- (length(h_off) - 1) %/% 2
v_off <- off$data$volts
t_off <- off$data$t_us / 1e6

cases <- list(
  list(label = "decimated, no filter", t = t_over[idx1], v = v_over[idx1],
       fs = fs_dec),
  list(label = "filtered, then decimated", t = t_filt[idx2], v = v_filt[idx2],
       fs = fs_dec),
  list(label = "already aliased, filtered",
       t = t_off[(lag_off + 1):(length(v_off) - lag_off)],
       v = fir_apply(v_off, h_off), fs = fs_off)
)

digital <- do.call(rbind, lapply(cases, function(cs) {
  a_sig <- sine_fit(cs$t, cs$v, F_SIGNAL)
  a_fold <- sine_fit(cs$t, cs$v, F_FOLDED)
  data.frame(case = cs$label, fs_Hz = cs$fs, n = length(cs$v),
             amp_signal_mV = a_sig$amplitude * 1e3,
             amp_folded_mV = a_fold$amplitude * 1e3,
             resid_mV = a_fold$resid_sd * 1e3)
}))
```

```{r}
#| label: out-digital
#| echo: false
#| output: asis
rows <- sprintf("| %s | %.0f | %d | %.2f | %.2f | %.3f |",
                digital$case, digital$fs_Hz, digital$n,
                digital$amp_signal_mV, digital$amp_folded_mV, digital$resid_mV)
cat(sprintf('
| Case | $f_s$ [Hz] | $n$ | Amplitude at %.0f Hz [mV] | Amplitude at %.0f Hz [mV] | Residual $\\sigma$ [mV] |
|---|---|---|---|---|---|
%s

: Filtering before and after the damage is done {#tbl-s3-digital}
', F_SIGNAL, F_FOLDED, paste(rows, collapse = "\n")))
```

```{r}
#| label: fig-s3-digital
#| echo: false
#| fig-width: 7.8
#| fig-height: 3.0
#| fig-cap: "Spectra of the three cases of @tbl-s3-digital. Filtering before decimation removes the interferer; decimating first folds it into the band, where it is indistinguishable from signal; filtering a record that is already aliased leaves it untouched."
spectra <- do.call(rbind, lapply(cases, function(cs) {
  x <- cs$v - mean(cs$v)
  n <- length(x)
  w <- 0.5 - 0.5 * cos(2 * pi * (0:(n - 1)) / (n - 1))
  mag <- Mod(fft(x * w)[1:(n %/% 2 + 1)]) * 2 / sum(w)
  data.frame(case = factor(cs$label, levels = vapply(cases, `[[`, character(1), "label")),
             freq = seq(0, cs$fs / 2, length.out = n %/% 2 + 1),
             mag = pmax(mag, 1e-7))
}))

ggplot(spectra, aes(freq, mag)) +
  geom_vline(xintercept = F_SIGNAL, linewidth = 0.3, linetype = "dotted",
             colour = "grey40") +
  geom_vline(xintercept = F_FOLDED, linewidth = 0.3, linetype = "dashed",
             colour = "grey70") +
  geom_line(linewidth = 0.25) +
  scale_y_log10() +
  facet_wrap(~case, scales = "free_x") +
  labs(x = "Frequency [Hz]", y = "Amplitude [V]") +
  theme(strip.text = element_text(size = 7))
```

```{r}
#| label: out-digital-text
#| echo: false
#| output: asis
naive <- digital[1, ]
clean <- digital[2, ]
aliased <- digital[3, ]
cat(sprintf('
Decimating without filtering puts %.1f mV at
%.0f Hz — the %.0f Hz interferer, folded, sitting
%.0f %% as large as the
signal and inside the band of interest. Filtering first leaves
%.2f mV there, a reduction of
%.0f dB.
The signal estimate moves by
%+.1f %%
between the two cases — part of that is the FIR\'s own droop at
%.0f Hz, and part is that the unfiltered estimate was biased by the
folded interferer sitting %.0f Hz away, too close to be
orthogonal to it over a record of this length. Neither part is noise, and
both are worth quantifying in your report.

The third row is the one to remember. The already-aliased recording, put
through the same kind of digital filter, still carries
%.1f mV at %.0f Hz — the alias is *in the
passband*, so the filter passes it, exactly as it passes a real signal there.
No digital filter can do anything else: the two are the same numbers.
', naive$amp_folded_mV, F_FOLDED, F_INTERFERER,
   naive$amp_folded_mV / naive$amp_signal_mV * 100,
   clean$amp_folded_mV,
   20 * log10(naive$amp_folded_mV / clean$amp_folded_mV),
   (clean$amp_signal_mV - naive$amp_signal_mV) / naive$amp_signal_mV * 100,
   F_SIGNAL, F_FOLDED - F_SIGNAL,
   aliased$amp_signal_mV, F_SIGNAL))
```

## Uncertainty budget

The measurand is the recovered generator amplitude of
[the handbook](handbook.qmd#eq-s3-recover-amplitude),
$\hat{V}_g = \hat{A}_\mathrm{adc} / (G\,\lvert H(f)\rvert)$, with the
[sensitivity coefficients](handbook.qmd#eq-s3-sensitivity) given there.

```{r}
#| label: budget
row_on <- pair[pair$case == "filtered", ]
a_adc <- row_on$amplitude_V
sigma_resid <- row_on$resid_V
n_samples <- row_on$n

h_mag <- filter_magnitude(F_GEN, FC, ORDER)
v_g <- a_adc / (GAIN * h_mag)

# gamma: the dimensionless factor in every derivative of |H(f)|
ratio2n <- (F_GEN / FC)^(2 * ORDER)
gamma <- ORDER * ratio2n / (1 + ratio2n)

c_amp <- 1 / (GAIN * h_mag)
c_gain <- -v_g / GAIN
c_fc <- -v_g * gamma / FC
c_f <- v_g * gamma / F_GEN

# The effective quantization step, by S1's method applied to this record.
enob_here <- enob(FSR, sigma_resid)
q_eff <- FSR / 2^enob_here
bin_width <- row_on$fs_Hz / n_samples

components <- list(
  list(quantity = "Amplitude estimation noise", value = a_adc, unit = "V",
       distribution = "type-a", std = sigma_resid * sqrt(2 / n_samples),
       sensitivity = c_amp),
  list(quantity = "Effective quantization", value = 0, unit = "V",
       distribution = "rectangular", half_width = q_eff / 2, sensitivity = c_amp),
  list(quantity = "Conditioning gain G", value = GAIN, unit = "V",
       distribution = "rectangular", half_width = U_GAIN_REL * GAIN,
       sensitivity = c_gain),
  list(quantity = "Conditioning offset", value = OFFSET, unit = "V",
       distribution = "rectangular", half_width = U_OFFSET_V, sensitivity = 0),
  list(quantity = "Filter corner frequency", value = FC, unit = "V",
       distribution = "rectangular", half_width = U_FC_REL * FC,
       sensitivity = c_fc),
  list(quantity = "Signal frequency", value = F_GEN, unit = "V",
       distribution = "rectangular", half_width = bin_width / 2,
       sensitivity = c_f)
)
budget <- budget_table(components, k = 2)
```

```{r}
#| label: out-budget
#| echo: false
#| output: asis
cat(budget_to_markdown(
  budget,
  caption = sprintf("Uncertainty budget for the recovered generator amplitude at $f = %.0f$ Hz", F_GEN),
  label = "tbl-s3-budget"))
```

```{r}
#| label: out-budget-text
#| echo: false
#| output: asis
dominant <- budget$table[which.max(budget$table$index_pct), ]
declared <- on$meta$channels[[1]]$reference_value
cat(sprintf('
Recovered generator amplitude $\\hat{V}_g = %.4f$ V against a setting
of %.4f V, with an expanded uncertainty $U = %.2f$ mV
at $k = 2$ — a departure of
%+.1f mV, or
%.1f times $U$.

The dominant component is **%s** at
%.0f %% of the combined variance. Note
$\\gamma = %.3f$: at %.0f Hz the filter is well into its
stopband, so a relative error in $f_c$ is amplified almost %.1f-fold
into the recovered amplitude. Calibrating the filter corner is worth more
here than any improvement to the converter.
', v_g, declared, budget$U * 1e3, (v_g - declared) * 1e3,
   abs(v_g - declared) / budget$U, dominant$quantity, dominant$index_pct,
   gamma, F_GEN, gamma))
```

::: {.callout-note}
## Why the offset row is there and equal to zero

The amplitude comes from a fit that carries its own mean level, so a DC
offset error does not reach it to first order — the sensitivity coefficient
is genuinely zero, not merely small. The row stays in the table because a
component checked and found not to matter is a result; a budget that silently
omits it looks like a budget that forgot.
:::

## What to report

1. @tbl-s3-conditioning: where the recordings sit in the converter's range,
   and how much headroom is left at each rail.
2. @tbl-s3-alias and @fig-s3-folding: predicted against observed apparent
   frequency, and the folding triangle. State how many of your generator
   settings are indistinguishable from each other in the recorded samples.
3. @tbl-s3-filter and @fig-s3-filter-pair: the measured attenuation against
   $\lvert H(f)\rvert$ at the true frequency, and why it is evaluated there
   and not at the apparent frequency.
4. @tbl-s3-design and @fig-s3-design: your own design case — your $f_s$, your
   passband edge, your required attenuation — with the order you would choose
   and the passband cost you accept.
5. @tbl-s3-digital and @fig-s3-digital: the three cases, and one sentence
   saying why the third cannot be fixed.
6. @tbl-s3-budget: the budget, the dominant component, and what you would
   calibrate first.
