---
title: "M2 — Analysis template (R)"
subtitle: "Hall-effect displacement measurement"
description: "A nonlinear calibration, a working range, the ratiometric correction and a stiffness cross-check, in R"
engine: knitr
toc: true
toc-depth: 3
code-fold: false
---

This template runs end to end on the reference data shipped with the
experiment. To analyse your own measurements, replace the files in `data/`
and change the paths in the first chunk — nothing else.

The Python template [`analysis-py.qmd`](analysis-py.qmd) has the same
sections, the same figure labels and the same output tables, and produces the
same numbers from the same data.

::: {.callout-note collapse="true"}
## Why the fit is written out rather than called from a library

The calibration curve is fitted by an explicit Gauss–Newton iteration, as
[the handbook explains](handbook.qmd#sec-m2-fit). `scipy.optimize.curve_fit`
and R's `nls` choose their own convergence criteria and do not reliably reach
the same parameters on the same data, and an analysis whose numbers depend on
which language ran it is not an analysis.
:::

## Setup

```{r}
#| label: setup
#| message: false
library(ggplot2)
library(patchwork)

source("../../shared/R/labtools.R")

theme_set(theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_line(linewidth = 0.25)))

CAL_RUNS <- sort(Sys.glob("data/cal-d*.csv"))
SUPPLY_RUN <- "data/supply-drift.csv"
CROSSCHECK_RUNS <- sort(Sys.glob("data/crosscheck-*.csv"))
M1_QUARTER_RUNS <- sort(Sys.glob("../M1-strain-load/data/bridge-quarter-*.csv"))

PIN_OUT <- 15         # Hall output
PIN_VCC <- 16         # supply-voltage sense
MAX_FIT_ITERATIONS <- 100
GRAVITY <- 9.80665    # m/s^2, the conventional standard value
```

```{r}
#| label: helpers
# The ratiometric reading r = V_out / V_cc, sample by sample. The sense
# channel is divided down by a known ratio, so the supply is recovered before
# the division. This is the whole point of the second channel: r does not
# depend on the supply, and V_out alone does.
ratio_of <- function(run) {
  wide <- pivot_channels(run$data, "volts")
  sense_gain <- NULL
  for (ch in run$meta$channels) if (ch$pin == PIN_VCC) sense_gain <- ch$conditioning$gain
  v_cc <- wide[[as.character(PIN_VCC)]] / sense_gain
  v_out <- wide[[as.character(PIN_OUT)]]
  list(r = v_out / v_cc, v_cc = v_cc, v_out = v_out)
}

# r(d) = 1/2 + A (1 + d/d0)^-n -- the ratiometric form, free of V_cc.
model_r <- function(d_mm, a, d0_mm, n) 0.5 + a / (1 + d_mm / d0_mm)^n

# Fits (A, d0) of the decay law by explicit damped Gauss-Newton. Written out,
# with the Jacobian in closed form, so that R and Python take the same
# arithmetic path and stop on the same criterion.
#
# The damping is not decoration: an undamped step from a plausible starting
# guess overshoots into negative d0 on this data and the Jacobian goes
# singular within a dozen iterations. The line search halves the step until d0
# stays positive and the sum of squares does not increase, which is
# deterministic and therefore reproduces identically in both languages.
gauss_newton <- function(d_mm, r_obs, a0, d0_0, n) {
  a <- a0
  d0 <- d0_0
  for (i in seq_len(MAX_FIT_ITERATIONS)) {
    u <- 1 + d_mm / d0
    f <- 0.5 + a * u^(-n)
    sse <- sum((r_obs - f)^2)
    j <- cbind(u^(-n), a * n * u^(-n - 1) * d_mm / d0^2)
    step <- solve(crossprod(j), crossprod(j, r_obs - f))

    lam <- 1
    while (lam > 1e-6) {
      a_try <- a + lam * step[1]
      d0_try <- d0 + lam * step[2]
      if (d0_try > 1e-6) {
        f_try <- 0.5 + a_try * (1 + d_mm / d0_try)^(-n)
        if (sum((r_obs - f_try)^2) <= sse) break
      }
      lam <- lam / 2
    }
    a <- a + lam * step[1]
    d0 <- d0 + lam * step[2]
    if (max(abs(lam * step[1]) / max(abs(a), 1e-12),
            abs(lam * step[2]) / max(abs(d0), 1e-12)) < 1e-12) break
  }

  u <- 1 + d_mm / d0
  resid <- r_obs - (0.5 + a * u^(-n))
  dof <- max(length(d_mm) - 2, 1)
  s_res <- sqrt(sum(resid^2) / dof)
  j <- cbind(u^(-n), a * n * u^(-n - 1) * d_mm / d0^2)
  cov <- solve(crossprod(j)) * s_res^2
  list(A = as.numeric(a), d0 = as.numeric(d0),
       u_A = sqrt(cov[1, 1]), u_d0 = sqrt(cov[2, 2]),
       s_res = s_res, residual = resid)
}

# d from r: the calibration curve read backwards.
invert <- function(r, a, d0, n) d0 * ((a / (r - 0.5))^(1 / n) - 1)

# dr/dd -- the sensitivity of the ratiometric reading, per mm.
d_r_d_d <- function(d_mm, a, d0, n) -a * n / d0 * (1 + d_mm / d0)^(-n - 1)

# Straight-line fit with the standard errors of both coefficients.
ols <- function(x, y) {
  n <- length(x)
  sxx <- sum((x - mean(x))^2)
  slope <- sum((x - mean(x)) * (y - mean(y))) / sxx
  intercept <- mean(y) - slope * mean(x)
  resid <- y - (intercept + slope * x)
  s <- sqrt(sum(resid^2) / max(n - 2, 1))
  list(slope = slope, intercept = intercept,
       u_slope = s / sqrt(sxx),
       u_intercept = s * sqrt(1 / n + mean(x)^2 / sxx),
       s_res = s)
}
```

```{r}
#| label: load
cal <- lapply(CAL_RUNS, read_run)
supply <- read_run(SUPPLY_RUN)
cross <- lapply(CROSSCHECK_RUNS, read_run)

board <- cal[[1]]$meta$board
FSR <- board$vref_mV / 1000
LSB <- lsb_volts(cal[[1]]$meta)
N_EXP <- cal[[1]]$meta$channels[[1]]$decay_exponent
TC_PER_K <- cal[[1]]$meta$channels[[1]]$sensor_temperature_coefficient_per_k
BEAM <- cross[[1]]$meta$beam
BASELINE_MM <- cross[[1]]$meta$channels[[1]]$baseline_distance_mm
```

```{r}
#| label: banner
#| echo: false
#| output: asis
cat(synthetic_banner(cal[[1]]$meta))
```

## Data quality

```{r}
#| label: quality
#| echo: false
#| output: asis
runs <- c(cal, list(supply), cross)
bad <- sum(vapply(runs, function(r) length(validate_meta(r$meta)) > 0, logical(1)))
lossy <- character(0)
for (r in runs) {
  st <- r$meta$stream_stats
  counters <- st[setdiff(names(st), "records")]
  if (any(vapply(counters, function(v) is.numeric(v) && v > 0, logical(1)))) {
    lossy <- c(lossy, r$meta$run)
  }
}
cat(sprintf('
%d runs read: %d calibration points, one supply-drift record
and %d beam cross-check points.
%s
%s
', length(runs), length(cal), length(cross),
   if (bad == 0) "All sidecars are complete."
   else sprintf("%d sidecar(s) have missing fields.", bad),
   if (length(lossy) == 0) "No run reports a lost record."
   else paste0("Runs with losses: ", paste(lossy, collapse = ", "))))
```

## Calibration fit

Each calibration run holds the magnet at a micrometer-set distance and records
both channels. The fit is on the **ratio**, not on the output voltage, so the
supply never enters.

```{r}
#| label: calibration
cal_df <- do.call(rbind, lapply(cal, function(run) {
  ch <- run$meta$channels[[1]]
  rr <- ratio_of(run)
  data.frame(run = run$meta$run,
             d_mm = ch$reference_value,
             u_d_mm = ch$reference_u,
             r_mean = mean(rr$r),
             r_sd = sd(rr$r),
             n = length(rr$r),
             v_out_mean = mean(rr$v_out),
             v_cc_mean = mean(rr$v_cc))
}))
cal_df <- cal_df[order(cal_df$d_mm), ]
rownames(cal_df) <- NULL

fit <- gauss_newton(cal_df$d_mm, cal_df$r_mean,
                    a0 = max(cal_df$r_mean) - 0.5,
                    d0_0 = median(cal_df$d_mm), n = N_EXP)
cal_df$residual <- fit$residual
```

```{r}
#| label: out-cal
#| echo: false
#| output: asis
rows <- sprintf("| %.0f | %.3f | %.6f | %.1f | %d | %+.1f |",
                cal_df$d_mm, cal_df$u_d_mm, cal_df$r_mean, cal_df$r_sd * 1e6,
                cal_df$n, cal_df$residual * 1e6)
cat(sprintf('
| $d$ [mm] | $u(d)$ [mm] | $r = V_\\mathrm{out}/V_\\mathrm{cc}$ | $s(r)$ [ppm] | $n$ | Residual [ppm] |
|---|---|---|---|---|---|
%s

: The calibration sweep and the residuals of the fitted decay law {#tbl-m2-cal}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: out-cal-text
#| echo: false
#| output: asis
cat(sprintf('
Fitted $A = %.5f$ with $u(A) = %.5f$, and
$d_0 = %.4f$ mm with $u(d_0) = %.4f$ mm, at a fixed
exponent $n = %d$. The residual standard deviation is
%.1f ppm of the ratio, and the largest single residual is
%.1f ppm.

Read the residual column of @tbl-m2-cal as a pattern, not as a number: scatter
means the two-parameter law is doing its job, while a systematic run of one
sign across the range would be evidence that the fixed exponent, or the
functional form itself, is wrong — and no amount of averaging repairs a model
of the wrong shape.
', fit$A, fit$u_A, fit$d0, fit$u_d0, N_EXP, fit$s_res * 1e6,
   max(abs(fit$residual)) * 1e6))
```

```{r}
#| label: fig-m2-cal
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.2
#| fig-cap: "The calibration curve. Top: the ratiometric reading against micrometer distance, with the fitted decay law. Bottom: residuals. The curve is steep and useful close in, and flattens into uselessness far out — which is what the next section quantifies."
grid <- seq(min(cal_df$d_mm), max(cal_df$d_mm), length.out = 300)
p_cal <- ggplot() +
  geom_line(data = data.frame(d = grid, r = model_r(grid, fit$A, fit$d0, N_EXP)),
            aes(d, r), linewidth = 0.5) +
  geom_point(data = cal_df, aes(d_mm, r_mean), size = 2) +
  labs(x = NULL, y = "r = V_out / V_cc")

p_res <- ggplot(cal_df, aes(d_mm, residual * 1e6)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.3) +
  geom_point(size = 2) +
  labs(x = "Micrometer distance d [mm]", y = "Residual [ppm]")

p_cal / p_res + plot_layout(heights = c(2, 1))
```

## Working range

Two criteria, pulling in opposite directions: resolution wants to be close to
the magnet, and linearity wants to be far from it.

```{r}
#| label: window
# The ADC's own noise, referred to the ratio, from the quietest calibration
# point's repeats.
sigma_r_adc <- min(cal_df$r_sd)

d_grid <- seq(min(cal_df$d_mm), max(cal_df$d_mm), length.out = 400)
sens <- abs(d_r_d_d(d_grid, fit$A, fit$d0, N_EXP))      # per mm
resolution_um <- sigma_r_adc / sens * 1e3               # micrometres
# Curvature, as the fractional change of sensitivity per millimetre.
curvature <- (N_EXP + 1) / (fit$d0 + d_grid) * 100      # % per mm

# The two criteria, written down before looking at the data. They pull in
# opposite directions, so the window is their overlap and not either alone;
# change them and the window moves, which is the point of stating them.
RESOLUTION_LIMIT_UM <- 40    # resolve 40 um or better
CURVATURE_LIMIT_PCT <- 30    # and no more than 30 % sensitivity change per mm
ok <- resolution_um <= RESOLUTION_LIMIT_UM & curvature <= CURVATURE_LIMIT_PCT
window <- if (any(ok)) range(d_grid[ok]) else c(NA_real_, NA_real_)
```

```{r}
#| label: fig-m2-window
#| echo: false
#| fig-width: 7.0
#| fig-height: 3.8
#| fig-cap: "Choosing a working range. Displacement resolution (left axis) improves close to the magnet; the curvature of the calibration curve (right axis) worsens there. The shaded band satisfies both stated criteria at once."
# The two quantities have different units, so the second is rescaled onto the
# first axis and given its own labelled secondary axis.
scale_factor <- max(log10(resolution_um)) / max(curvature)
plot_df <- data.frame(
  d = rep(d_grid, 2),
  value = c(log10(resolution_um), curvature * scale_factor),
  what = rep(c("resolution [µm]", "curvature [% per mm]"), each = length(d_grid)))

p <- ggplot(plot_df, aes(d, value, colour = what)) +
  geom_line(linewidth = 0.5) +
  scale_y_continuous(name = "log10 resolution [µm]",
                     sec.axis = sec_axis(~ . / scale_factor,
                                         name = "Sensitivity change [% per mm]")) +
  geom_hline(yintercept = log10(RESOLUTION_LIMIT_UM), linewidth = 0.3,
             linetype = "dashed", colour = "grey40") +
  geom_hline(yintercept = CURVATURE_LIMIT_PCT * scale_factor, linewidth = 0.3,
             linetype = "dotted", colour = "grey40") +
  labs(x = "Distance d [mm]", colour = NULL) +
  theme(legend.position = "top")
if (any(ok)) {
  p <- p + annotate("rect", xmin = window[1], xmax = window[2],
                    ymin = -Inf, ymax = Inf, fill = "grey85", alpha = 0.5)
}
p
```

```{r}
#| label: out-window
#| echo: false
#| output: asis
d_work <- BASELINE_MM
sens_work <- abs(d_r_d_d(d_work, fit$A, fit$d0, N_EXP))
cat(sprintf('
| Quantity | Value |
|---|---|
| Noise on $r$, from repeats | %.1f ppm |
| Resolution criterion | ≤ %.0f µm |
| Curvature criterion | ≤ %.0f %% per mm |
| Usable window | %.1f – %.1f mm |
| Chosen working point | %.1f mm |
| Sensitivity there | %.1f ppm per µm |
| Resolution there | %.2f µm |

: The working range, from two stated criteria {#tbl-m2-window}
', sigma_r_adc * 1e6, RESOLUTION_LIMIT_UM, CURVATURE_LIMIT_PCT,
   window[1], window[2], d_work, sens_work * 1e6,
   sigma_r_adc / sens_work * 1e3))
```

## Ratiometric correction

The supply is deliberately drifted while the magnet is held still. Nothing
moves; only the rail does.

```{r}
#| label: supply
ch_supply <- supply$meta$channels[[2]]
sup <- ratio_of(supply)
t_s <- sort(unique(supply$data$t_us)) / 1e6
t_s <- t_s - t_s[1]

# Ratiometric: invert the calibration curve at the measured ratio.
d_ratio <- invert(sup$r, fit$A, fit$d0, N_EXP)
# Absolute: assume the supply is at its nominal value and never checked.
v_cc_nominal <- ch_supply$reference_value
d_absolute <- invert(sup$v_out / v_cc_nominal, fit$A, fit$d0, N_EXP)

drift_pct <- (sup$v_cc[length(sup$v_cc)] / sup$v_cc[1] - 1) * 100

# The *trend* over the record, not its span: the span of any noisy record is
# about 8 sigma whatever the supply does, which would hide the effect this
# section exists to show. A straight-line fit separates drift from noise.
trend_abs <- ols(t_s, d_absolute)
trend_ratio <- ols(t_s, d_ratio)
duration_s <- t_s[length(t_s)] - t_s[1]
drift_abs_um <- trend_abs$slope * duration_s * 1e3
drift_ratio_um <- trend_ratio$slope * duration_s * 1e3
u_drift_abs_um <- trend_abs$u_slope * duration_s * 1e3
u_drift_ratio_um <- trend_ratio$u_slope * duration_s * 1e3
```

```{r}
#| label: fig-m2-supply
#| echo: false
#| fig-width: 7.0
#| fig-height: 4.6
#| fig-cap: "The magnet is stationary throughout. Top: the measured supply. Bottom: the displacement each reading implies — absolute, which tracks the rail, and ratiometric, which does not."
p_vcc <- ggplot(data.frame(t = t_s, v = sup$v_cc), aes(t, v)) +
  geom_line(linewidth = 0.3) +
  labs(x = NULL, y = "Vcc [V]")

move <- rbind(
  data.frame(t = t_s, um = (d_absolute - d_absolute[1]) * 1e3,
             what = "absolute reading"),
  data.frame(t = t_s, um = (d_ratio - d_ratio[1]) * 1e3,
             what = "ratiometric"))
p_move <- ggplot(move, aes(t, um, colour = what)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.3) +
  labs(x = "Time [s]", y = "Apparent movement [µm]", colour = NULL) +
  theme(legend.position = "top")

p_vcc / p_move + plot_layout(heights = c(1, 2))
```

```{r}
#| label: out-supply
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Absolute reading | Ratiometric reading |
|---|---|---|
| Supply drift over the record | %+.2f %% | %+.2f %% |
| Fitted apparent movement | %+.1f µm | %+.3f µm |
| Its own standard uncertainty | %.2f µm | %.3f µm |
| Resolved? | %s | %s |

: What the second channel buys {#tbl-m2-supply}
', drift_pct, drift_pct, drift_abs_um, drift_ratio_um,
   u_drift_abs_um, u_drift_ratio_um,
   if (abs(drift_abs_um) > 2 * u_drift_abs_um) "yes" else "no",
   if (abs(drift_ratio_um) > 2 * u_drift_ratio_um) "yes" else "no"))
```

```{r}
#| label: out-supply-text
#| echo: false
#| output: asis
cat(sprintf('
A supply droop of %.2f %% fakes %.0f µm of
movement if the rail is assumed rather than measured — around
%.0f times the uncertainty of the trend
itself, so it would not look like noise, it would look like a measurement.
Dividing by the *measured* supply leaves %+.3f µm, against an
uncertainty of %.3f µm on that number: the apparent movement
is no longer distinguishable from zero. The error is gone, not reduced.

This is why the supply does not appear as a row in the budget below. One extra
analog channel removed an entire uncertainty component, which is a better
trade than any amount of post-processing.
', abs(drift_pct), abs(drift_abs_um), abs(drift_abs_um) / u_drift_abs_um,
   drift_ratio_um, u_drift_ratio_um))
```

## Stiffness cross-check

The same reference masses that M1 weighed with strain gauges, measured again
as a tip displacement. Two independent routes to the beam's stiffness.

```{r}
#| label: crosscheck
cc <- do.call(rbind, lapply(cross, function(run) {
  ch <- run$meta$channels[[1]]
  rr <- ratio_of(run)
  d_mm <- invert(mean(rr$r), fit$A, fit$d0, N_EXP)
  data.frame(run = run$meta$run, force_N = ch$reference_value,
             u_force_N = ch$reference_u, d_mm = d_mm,
             delta_mm = BASELINE_MM - d_mm, r_sd = sd(rr$r), n = length(rr$r))
}))
cc <- cc[order(cc$force_N), ]
rownames(cc) <- NULL
fit_m2 <- ols(cc$force_N, cc$delta_mm / 1e3)     # m per N

L <- BEAM$length_m; B <- BEAM$width_m; H <- BEAM$thickness_m
XG <- BEAM$gauge_position_m
I_BEAM <- B * H^3 / 12
E_M2 <- L^3 / (3 * I_BEAM * fit_m2$slope)
U_E_M2 <- E_M2 * fit_m2$u_slope / fit_m2$slope

m1 <- do.call(rbind, lapply(M1_QUARTER_RUNS, function(path) {
  run <- read_run(path)
  ch <- run$meta$channels[[1]]
  data.frame(force_N = ch$reference_value,
             v_mean = mean(run$data$volts),
             gain = ch$conditioning$gain,
             offset = ch$conditioning$offset,
             k = ch$gauge_factor)
}))
m1 <- m1[order(m1$force_N), ]
rownames(m1) <- NULL
```

```{r}
#| label: crosscheck-m1
# Strain from the quarter-bridge reading: V_adc = G (k eps Vex / 4) + offset.
m1_meta <- read_run(M1_QUARTER_RUNS[1])$meta
EXCITATION <- if (!is.null(m1_meta$channels[[1]]$excitation_v)) {
  m1_meta$channels[[1]]$excitation_v
} else if (!is.null(cal[[1]]$meta$excitation_v)) {
  cal[[1]]$meta$excitation_v
} else 3.3

m1$strain <- 4 * (m1$v_mean - m1$offset) / (m1$gain * m1$k * EXCITATION)
fit_m1 <- ols(m1$force_N, m1$strain)
E_M1 <- 6 * (L - XG) / (B * H^2 * fit_m1$slope)
U_E_M1 <- E_M1 * fit_m1$u_slope / fit_m1$slope

ratio_e <- E_M1 / E_M2
u_ratio_e <- ratio_e * sqrt((U_E_M1 / E_M1)^2 + (U_E_M2 / E_M2)^2)
```

```{r}
#| label: out-crosscheck
#| echo: false
#| output: asis
rows <- sprintf("| %.3f | %.4f | %.4f | %.1f |",
                cc$force_N, cc$u_force_N, cc$d_mm, cc$delta_mm * 1e3)
cat(sprintf('
| Force [N] | $u(F)$ [N] | Distance $d$ [mm] | Tip displacement [µm] |
|---|---|---|---|
%s

: The beam cross-check, read through the calibration curve {#tbl-m2-crosscheck}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: fig-m2-crosscheck
#| echo: false
#| fig-width: 7.0
#| fig-height: 3.4
#| fig-cap: "Tip displacement against applied force, with the straight-line fit whose slope is the beam's compliance. The intercept is the zero of the Hall reading, not a physical offset."
grid <- seq(0, max(cc$force_N) * 1.05, length.out = 100)
ggplot() +
  geom_line(data = data.frame(f = grid,
                              um = (fit_m2$intercept + fit_m2$slope * grid) * 1e6),
            aes(f, um), linewidth = 0.45) +
  geom_point(data = cc, aes(force_N, delta_mm * 1e3), size = 2) +
  labs(x = "Applied force [N]", y = "Tip displacement [µm]")
```

```{r}
#| label: out-crosscheck-text
#| echo: false
#| output: asis
cat(sprintf('
| Route | Fitted slope | Young\'s modulus |
|---|---|---|
| M1, strain per force | %.4f µstrain/N | %.2f ± %.2f GPa |
| M2, displacement per force | %.4f µm/N | %.2f ± %.2f GPa |
| Ratio $E_\\mathrm{M1}/E_\\mathrm{M2}$ | — | %.4f ± %.4f |

: Two independent routes to the same beam {#tbl-m2-stiffness}

The **ratio** is the result. Both routes need the beam\'s geometry, which is a
placeholder here, so neither modulus is a measurement of anything — but the
same placeholder geometry enters both, so a ratio near 1 says the two
mechanical models agree with each other, and a ratio away from 1 is evidence
about the clamp rather than about the placeholders. This one is
%.2f %% from unity, against a combined uncertainty of
%.2f %%.
', fit_m1$slope * 1e6, E_M1 / 1e9, U_E_M1 / 1e9,
   fit_m2$slope * 1e6, E_M2 / 1e9, U_E_M2 / 1e9,
   ratio_e, u_ratio_e, abs(ratio_e - 1) * 100, u_ratio_e * 100))
```

## Uncertainty budget

The measurand is the measured tip displacement at the largest reference force,
recovered by [inverting the calibration curve](handbook.qmd#eq-m2-recover)
with the [sensitivity coefficients](handbook.qmd#eq-m2-sensitivity) given
there.

```{r}
#| label: budget
last <- cc[nrow(cc), ]
r_op <- model_r(last$d_mm, fit$A, fit$d0, N_EXP)
d_hat <- invert(r_op, fit$A, fit$d0, N_EXP)

# Sensitivity coefficients of eq-m2-recover, in mm per unit of each input.
u_arg <- fit$A / (r_op - 0.5)
c_r <- -fit$d0 / N_EXP * fit$A * u_arg^(1 / N_EXP - 1) / (r_op - 0.5)^2
c_A <- fit$d0 / N_EXP * u_arg^(1 / N_EXP - 1) / (r_op - 0.5)
c_d0 <- d_hat / fit$d0

n_op <- last$n
sigma_r_op <- last$r_sd

# Effective quantization, by S1's method applied to this record, referred to r.
vcc_ref <- mean(vapply(supply$meta$channels,
                       function(ch) if (ch$pin == PIN_VCC) ch$reference_value
                                    else NA_real_, numeric(1)), na.rm = TRUE)
enob_here <- enob(FSR, sigma_r_op * vcc_ref)
q_eff_r <- FSR / 2^enob_here / (board$vref_mV / 1000)

DELTA_T_K <- 5    # the bench temperature excursion the budget bounds

components <- list(
  list(quantity = "Repeatability of r", value = r_op, unit = "mm",
       distribution = "type-a", std = sigma_r_op / sqrt(n_op), sensitivity = c_r),
  list(quantity = "Effective quantization", value = 0, unit = "mm",
       distribution = "rectangular", half_width = q_eff_r / 2, sensitivity = c_r),
  list(quantity = "Calibration amplitude A", value = fit$A, unit = "mm",
       distribution = "type-a", std = fit$u_A, sensitivity = c_A),
  list(quantity = "Calibration distance d0", value = fit$d0, unit = "mm",
       distribution = "type-a", std = fit$u_d0, sensitivity = c_d0),
  list(quantity = "Micrometer reference", value = 0, unit = "mm",
       distribution = "rectangular", half_width = max(cal_df$u_d_mm),
       sensitivity = 1),
  list(quantity = "Sensor temperature coefficient", value = 0, unit = "mm",
       distribution = "rectangular",
       half_width = TC_PER_K * DELTA_T_K * fit$A, sensitivity = c_A)
)
budget <- budget_table(components, k = 2)
```

```{r}
#| label: out-budget
#| echo: false
#| output: asis
cat(budget_to_markdown(
  budget,
  caption = sprintf("Uncertainty budget for the tip displacement at $F = %.3f$ N",
                    last$force_N),
  label = "tbl-m2-budget"))
```

```{r}
#| label: out-budget-text
#| echo: false
#| output: asis
dominant <- budget$table[which.max(budget$table$index_pct), ]
cat(sprintf('
Measured tip displacement %.1f µm with an expanded
uncertainty $U = %.1f$ µm at $k = 2$ — a relative uncertainty
of %.1f %%.

The dominant component is **%s** at
%.0f %% of the combined variance. Note what is *not*
in the table: the supply voltage, because the reading is ratiometric. An
absolute reading would carry a row of its own, and @tbl-m2-supply says how
large it would be.

The effective-quantization row is deliberately conservative. It stands for the
part of the converter\'s noise that does **not** average away with $n$, and it
is written from the noise itself rather than from the nominal code width, in
the same way as in
[S1\'s budget](../S1-daq-static/handbook.qmd#sec-s1-budget). Whether that floor
is real, and where averaging genuinely stops helping, is exactly the question
[T3](../T3-long-term-stability/index.qmd) answers with an Allan deviation; a
group that has done T3 should replace this bound with the measured one and say
what changed.
', last$delta_mm * 1e3, budget$U * 1e3,
   budget$U / last$delta_mm * 100, dominant$quantity, dominant$index_pct))
```

## What to report

1. @tbl-m2-cal and @fig-m2-cal: the fitted curve and its residuals, with a
   statement about whether the fixed exponent is earning its keep.
2. @fig-m2-window and @tbl-m2-window: your two criteria, written down as
   numbers before you look at the data, and the window they imply.
3. @tbl-m2-supply and @fig-m2-supply: what the second channel is worth, in
   micrometres.
4. @tbl-m2-crosscheck, @fig-m2-crosscheck and @tbl-m2-stiffness: the two
   routes to the beam's stiffness and their ratio, with the ratio — not the
   moduli — presented as the result.
5. @tbl-m2-budget: the budget, the dominant component, and one sentence on
   why the supply is absent from it.
