---
title: "S1 — Analysis template (R)"
subtitle: "Static characterization of the DAQ"
description: "Offset, gain, noise, effective bits and timestamp quality, in R"
engine: knitr
toc: true
toc-depth: 3
code-fold: false
---

This template runs end to end on the reference data shipped with the
experiment, so you can read the whole analysis before you touch a board. To
analyse your own measurements, replace the files in `data/` and change the
paths in the first chunk — nothing else.

The Python template [`analysis-py.qmd`](analysis-py.qmd) has the same
sections, the same figure labels and the same output tables, and produces the
same numbers from the same data.

::: {.callout-note collapse="true"}
## How the chunks are organised

Chunks that compute something show their code: they are the part you edit.
Chunks that only format a result into a table are hidden, because reading them
teaches nothing about measurement. Every table is emitted as markdown from a
chunk with `#| output: asis`, which is what lets the same table appear in the
HTML page and in the PDF handbook.
:::

## Setup

```{r}
#| label: setup
#| message: false
library(ggplot2)
library(purrr)
library(patchwork)

source("../../shared/R/labtools.R")

# The five runs of the procedure. Point these at your own files.
CAL_RUNS <- c("data/cal-0v.csv", "data/cal-1v25.csv", "data/cal-2v50.csv")
NOISE_RUNS <- c(linux = "data/noise-linux.csv", macos = "data/noise-macos.csv")
PRIMARY <- "linux"        # the host whose run carries the main analysis
PIN_REFERENCE <- 15       # the pin fed by the voltage reference
PIN_SHORTED <- 16         # the pin shorted to GND
```

```{r}
#| label: load
runs <- map(NOISE_RUNS, read_run)
primary <- runs[[PRIMARY]]
board <- primary$meta$board

LSB <- lsb_volts(primary$meta)          # volts per code, 2^n - 1 full scale
FSR <- board$vref_mV / 1000             # full-scale range, volts
N_BITS <- board$adc_bits
```

```{r}
#| label: banner
#| echo: false
#| output: asis
cat(synthetic_banner(primary$meta))
```

## Data quality

Before any number is computed, the metadata must be complete and the recording
must be intact. A run with lost records is not a slow run: it is a run with
holes in its time axis.

```{r}
#| label: quality
#| echo: false
#| output: asis
for (name in names(runs)) {
  meta <- runs[[name]]$meta
  problems <- validate_meta(meta)
  stats <- meta$stream_stats
  counters <- stats[setdiff(names(stats), "records")]
  lossy <- counters[map_lgl(counters, ~ is.numeric(.x) && .x > 0)]
  cat("\n**", name, "** — ", stats$records, " records; ",
      if (length(problems) == 0) "metadata complete"
      else paste0("metadata problems: ", paste(problems, collapse = "; ")),
      "; ",
      if (length(lossy) == 0) "no losses reported"
      else paste0("losses: ",
                  paste(names(lossy), unlist(lossy), sep = " = ", collapse = ", ")),
      "\n", sep = "")
}
```

```{r}
#| label: out-board
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| Board | `%s` |
| Reported ADC resolution | %d bit |
| Reference voltage $V_\\mathrm{ref}$ | %.3f V |
| One LSB, $q = V_\\mathrm{ref}/(2^{n}-1)$ | %.2f µV |
| Quantization uncertainty $u_q = q/\\sqrt{12}$ | %.2f µV |

: What the board says about itself {#tbl-s1-board}
', board$model, N_BITS, FSR, LSB * 1e6, u_quantization(LSB) * 1e6))
```

## Static calibration

Each calibration run is a few hundred readings of a voltage measured
independently with the DMM. The mean indication at each point, fitted against
the reference value, gives the offset and the gain error of the
[calibration model](handbook.qmd#eq-s1-fit) in the handbook.

```{r}
#| label: calibration
points <- map(CAL_RUNS, function(path) {
  run <- read_run(path)
  pin <- run$meta$acquisition$pins[[1]]
  channel <- Filter(function(c) c$pin == pin, run$meta$channels)[[1]]
  volts <- run$data$volts[run$data$pin == pin]
  data.frame(
    run = run$meta$run,
    reference_V = channel$reference_value,
    u_reference_V = channel$reference_u,
    mean_indicated_V = mean(volts),
    sd_V = sd(volts),
    n = length(volts)
  )
})
cal <- list_rbind(points)
cal <- cal[order(cal$reference_V), ]
rownames(cal) <- NULL

# Ordinary least squares of the indication on the reference value.
x <- cal$reference_V
y <- cal$mean_indicated_V
n_pts <- length(x)
sxx <- sum((x - mean(x))^2)
b <- sum((x - mean(x)) * (y - mean(y))) / sxx
a <- mean(y) - b * mean(x)
cal$residual_V <- y - (a + b * x)

# Standard errors of the fit, from the residuals (n - 2 degrees of freedom).
s_res <- sqrt(sum(cal$residual_V^2) / (n_pts - 2))
u_b <- s_res / sqrt(sxx)
u_a <- s_res * sqrt(1 / n_pts + mean(x)^2 / sxx)
```

```{r}
#| label: out-cal
#| echo: false
#| output: asis
rows <- sprintf("| %s | %.4f | %.5f | %.6f | %.6f | %d | %+.1f |",
                cal$run, cal$reference_V, cal$u_reference_V,
                cal$mean_indicated_V, cal$sd_V, cal$n, cal$residual_V * 1e6)
cat(sprintf('
| Run | Reference (V) | $u$(ref) (V) | Mean indication (V) | $s$ (V) | $n$ | Residual (µV) |
|---|---|---|---|---|---|---|
%s

: Calibration points and fit residuals {#tbl-s1-cal}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: out-cal-text
#| echo: false
#| output: asis
cat(sprintf('
Fitted offset $a = %.3f$ mV with $u(a) = %.3f$ mV; fitted
slope $b = %.6f$, that is a gain error of $%+.3f\\,\\%%$, with
$u(b) = %.6f$. The largest residual is
$%.1f$ µV, or
$%.1f$ LSB — the bound that enters the
budget as residual nonlinearity, because three points cannot correct it.
', a * 1e3, u_a * 1e3, b, (b - 1) * 100, u_b,
   max(abs(cal$residual_V)) * 1e6, max(abs(cal$residual_V)) / LSB))
```

```{r}
#| label: fig-s1-cal
#| echo: false
#| fig-width: 7.0
#| fig-height: 5.2
#| fig-cap: "Calibration: indication against reference (top) and fit residuals (bottom). Three points give one degree of freedom — enough to see nonlinearity, not enough to correct it."
p_fit <- ggplot(cal, aes(reference_V, mean_indicated_V)) +
  geom_abline(intercept = a, slope = b, linewidth = 0.5) +
  geom_point(size = 2.4) +
  labs(x = NULL, y = "Indicated voltage (V)",
       subtitle = sprintf("V_ind = %.2f mV + %.5f V", a * 1e3, b))

p_res <- ggplot(cal, aes(reference_V, residual_V * 1e6)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.4) +
  geom_point(size = 2.4) +
  labs(x = "Reference voltage (V)", y = "Residual (µV)")

p_fit / p_res + plot_layout(heights = c(2, 1))
```

## Noise floor

The long run carries two channels: the reference, and an input shorted to
ground. The shorted channel is the floor — whatever appears there is the
board's, not the source's.

```{r}
#| label: noise
wide <- pivot_channels(primary$data, "volts")
v_ref <- wide[[as.character(PIN_REFERENCE)]]
v_short <- wide[[as.character(PIN_SHORTED)]]

noise <- data.frame(
  channel = c("reference", "shorted"),
  pin = c(PIN_REFERENCE, PIN_SHORTED),
  mean_V = c(mean(v_ref), mean(v_short)),
  sd_V = c(sd(v_ref), sd(v_short)),
  n = c(length(v_ref), length(v_short))
)
noise$sd_LSB <- noise$sd_V / LSB
noise$peak_to_peak_LSB <- c(diff(range(v_ref)), diff(range(v_short))) / LSB
noise$u_mean_V <- noise$sd_V / sqrt(noise$n)
```

```{r}
#| label: out-noise
#| echo: false
#| output: asis
rows <- sprintf("| %s | %d | %.6f | %.6f | %.1f | %.0f | %.2f |",
                noise$channel, noise$pin, noise$mean_V, noise$sd_V,
                noise$sd_LSB, noise$peak_to_peak_LSB, noise$u_mean_V * 1e6)
cat(sprintf('
| Channel | Pin | Mean (V) | $\\sigma$ (V) | $\\sigma$ (LSB) | Peak-to-peak (LSB) | $\\sigma/\\sqrt{n}$ (µV) |
|---|---|---|---|---|---|---|
%s

: Noise on the two channels of the `%s` run, $n = %d$ {#tbl-s1-noise}
', paste(rows, collapse = "\n"), PRIMARY, nrow(wide)))
```

```{r}
#| label: fig-s1-hist
#| echo: false
#| fig-width: 7.4
#| fig-height: 3.4
#| fig-cap: "Distribution of the readings on each channel, in ADC codes relative to the channel mean, with a Gaussian of the same standard deviation. A converter limited by quantization alone would put every reading in one or two bins."
hist_data <- map(
  list(c("reference", PIN_REFERENCE), c("shorted", PIN_SHORTED)),
  function(spec) {
    pin <- as.integer(spec[2])
    codes <- primary$data$raw[primary$data$pin == pin]
    centred <- codes - round(mean(codes))
    data.frame(panel = sprintf("%s (pin %d), sigma = %.1f LSB",
                               spec[1], pin, sd(centred)),
               centred = centred)
  }) |> list_rbind()

gaussians <- map(split(hist_data, hist_data$panel), function(d) {
  grid <- seq(min(d$centred), max(d$centred), length.out = 200)
  data.frame(panel = d$panel[1], centred = grid,
             density = dnorm(grid, 0, sd(d$centred)))
}) |> list_rbind()

ggplot(hist_data, aes(centred)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 1, alpha = 0.75,
                 fill = "#2a7fa8") +
  geom_line(data = gaussians, aes(y = density), linewidth = 0.5, colour = "grey20") +
  facet_wrap(~panel, scales = "free_x") +
  labs(x = "Code − mean code (LSB)", y = "Density")
```

## Effective number of bits

The [ENOB definition](handbook.qmd#eq-s1-enob) turns the measured spread into
the resolution an ideal converter would need to produce it. This is the
*static* effective resolution: measured on a constant input, it says nothing
about behaviour at speed, which is what S2 measures.

```{r}
#| label: effective-bits
sigma_ref <- noise$sd_V[noise$channel == "reference"]
sigma_short <- noise$sd_V[noise$channel == "shorted"]

enob_ref <- enob(FSR, sigma_ref)
enob_short <- enob(FSR, sigma_short)
enob_nf <- log2(FSR / (6.6 * sigma_ref))    # noise-free, +-3.3 sigma span
q_eff <- FSR / 2^enob_ref                   # effective quantization step
```

```{r}
#| label: out-enob
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Reference channel | Shorted channel |
|---|---|---|
| Nominal resolution | %d bit | %d bit |
| Effective number of bits | %.2f bit | %.2f bit |
| Noise-free resolution | %.2f bit | — |
| Effective quantization step $q_\\mathrm{eff}$ | %.1f µV | — |
| Nominal step $q$ | %.2f µV | %.2f µV |

: Effective resolution against the nominal one {#tbl-s1-enob}
', N_BITS, N_BITS, enob_ref, enob_short, enob_nf, q_eff * 1e6,
   LSB * 1e6, LSB * 1e6))
```

```{r}
#| label: out-enob-text
#| echo: false
#| output: asis
cat(sprintf('
The board reports **%d bits** and delivers **%.1f**. The
effective quantization step is %.0f times the nominal one, and it
is that number, not $q$, that belongs in the uncertainty budget below.
', N_BITS, enob_ref, q_eff / LSB))
```

## Timestamps

Every record carries the device's own `micros()`. Fitting the
[sampling grid](handbook.qmd#eq-s1-grid) and reading the residuals says
whether the sampling was regular, and if not, how it failed.

```{r}
#| label: timing
timing <- map(runs, function(run) timing_summary(run$data))
primary_timing <- timing[[PRIMARY]]
ts <- primary_timing$stats
```

```{r}
#| label: fig-s1-timing
#| echo: false
#| fig-width: 7.6
#| fig-height: 3.4
#| fig-cap: "Sampling grid of the primary run. Left: the interval between consecutive records. Right: the residuals of the least-squares grid fit. A flat band is bounded jitter, a slope is a wrong period, a step is a schedule restart."
p_int <- ggplot(data.frame(i = seq_along(primary_timing$intervals_us),
                           d = primary_timing$intervals_us),
                aes(i, d)) +
  geom_line(linewidth = 0.2) +
  geom_hline(yintercept = ts$median_interval_us, linetype = "dashed",
             linewidth = 0.3, colour = "grey30") +
  labs(x = "Record index", y = "Interval (µs)",
       subtitle = sprintf("median %.0f µs", ts$median_interval_us))

p_res <- ggplot(data.frame(i = seq_along(primary_timing$residuals_us),
                           r = primary_timing$residuals_us),
                aes(i, r)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.25) +
  labs(x = "Record index", y = "Grid-fit residual (µs)")

p_int | p_res
```

```{r}
#| label: out-timing
#| echo: false
#| output: asis
cat(sprintf('
| Quantity | Value |
|---|---|
| Requested period | %d µs |
| Fitted period $\\hat{T}$ | %.2f µs |
| Achieved rate | %.2f Hz |
| Interval: median / min / max | %.0f / %.0f / %.0f µs |
| Interval standard deviation | %.1f µs |
| Residual RMS | %.1f µs |
| Largest absolute residual | %.1f µs |
| Intervals longer than $1.5\\times$ the median | %d |
| Timestamp quantization $u_q$ | %.2f µs |

: Sampling grid of the `%s` run {#tbl-s1-timing}
', primary$meta$acquisition$requested_period_us, ts$period_us,
   ts$achieved_rate_hz, ts$median_interval_us, ts$min_interval_us,
   ts$max_interval_us, ts$sd_interval_us, ts$residual_rms_us,
   ts$residual_max_abs_us, ts$n_long_intervals, u_quantization(1), PRIMARY))
```

## The host as an influence quantity

The same board, the same firmware and the same request, recorded on two
different hosts. Nothing about the measurand changed.

```{r}
#| label: out-hosts
#| echo: false
#| output: asis
rows <- map_chr(names(runs), function(name) {
  s <- timing[[name]]$stats
  meta <- runs[[name]]$meta
  sprintf("| %s | `%s` | %.2f | %.2f | %.1f | %.1f | %d | %s |",
          meta$host$os, meta$host$machine, s$achieved_rate_hz, s$period_us,
          s$sd_interval_us, s$residual_rms_us, s$n_long_intervals,
          as.character(meta$stream_stats$device_overruns))
})
cat(sprintf('
| Host | Machine | Achieved (Hz) | $\\hat{T}$ (µs) | $s$(interval) (µs) | Residual RMS (µs) | Long intervals | Device overruns |
|---|---|---|---|---|---|---|---|
%s

: The same acquisition on two hosts {#tbl-s1-hosts}
', paste(rows, collapse = "\n")))
```

```{r}
#| label: fig-s1-hosts
#| echo: false
#| fig-width: 7.0
#| fig-height: 3.2
#| fig-cap: "Grid-fit residuals for both hosts on the same axes. The board is identical in the two runs; what differs is how often the host was late to drain the endpoint."
host_residuals <- map(names(runs), function(name) {
  data.frame(host = runs[[name]]$meta$host$os,
             i = seq_along(timing[[name]]$residuals_us),
             r = timing[[name]]$residuals_us)
}) |> list_rbind()

ggplot(host_residuals, aes(i, r, colour = host)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
  geom_line(linewidth = 0.25) +
  labs(x = "Record index", y = "Grid-fit residual (µs)", colour = NULL)
```

## Uncertainty budget

The measurand is the [corrected voltage](handbook.qmd#eq-s1-correct)
$\hat{V} = (V_\mathrm{ind} - a)/b$, evaluated at the reference point, with the
[sensitivity coefficients](handbook.qmd#eq-s1-sensitivity) of the handbook.

```{r}
#| label: budget
v_ind <- noise$mean_V[noise$channel == "reference"]
n_avg <- noise$n[noise$channel == "reference"]
v_hat <- (v_ind - a) / b

c_vind <- 1 / b
c_a <- -1 / b
c_b <- -(v_ind - a) / b^2

u_dmm <- max(cal$u_reference_V)
inl_bound <- max(abs(cal$residual_V))

components <- list(
  list(quantity = "Repeatability of the mean", value = v_ind, unit = "V",
       distribution = "type-a", std = sigma_ref / sqrt(n_avg),
       sensitivity = c_vind),
  list(quantity = "Effective quantization", value = 0, unit = "V",
       distribution = "rectangular", half_width = q_eff / 2, sensitivity = c_vind),
  list(quantity = "Offset a from the fit", value = a, unit = "V",
       distribution = "type-a", std = u_a, sensitivity = c_a),
  list(quantity = "Slope b from the fit", value = b, unit = "V",
       distribution = "type-a", std = u_b, sensitivity = c_b),
  list(quantity = "Reference value (DMM)", value = 0, unit = "V",
       distribution = "type-a", std = u_dmm, sensitivity = 1),
  list(quantity = "Residual nonlinearity", value = 0, unit = "V",
       distribution = "rectangular", half_width = inl_bound, sensitivity = c_vind)
)
budget <- budget_table(components, k = 2)
```

```{r}
#| label: out-budget
#| echo: false
#| output: asis
cat(budget_to_markdown(
  budget,
  caption = sprintf("Uncertainty budget for the corrected reading at $V_\\mathrm{ind} = %.4f$ V", v_ind),
  label = "tbl-s1-budget"))
```

```{r}
#| label: out-budget-text
#| echo: false
#| output: asis
dominant <- budget$table[which.max(budget$table$index_pct), ]
quant <- budget$table$index_pct[budget$table$quantity == "Effective quantization"]
cat(sprintf('
Corrected value $\\hat{V} = %.5f$ V, against a reference value of
%.4f V, with an expanded uncertainty
$U = %.3f$ mV at $k = 2$. The dominant component is
**%s** at %.0f %% of the combined
variance; the effective quantization contributes %.1f %%.
', v_hat, max(cal$reference_V), budget$U * 1e3, dominant$quantity,
   dominant$index_pct, quant))
```

::: {.callout-note}
## What this budget assumes

The components are treated as uncorrelated, which $a$ and $b$ are not: they
come from the same fit. The correlation matters most at the ends of the
calibrated range, and propagating the covariance matrix of the fit instead is
the natural extension (see
[Calibrating offset and gain](handbook.qmd#sec-s1-calibration)). The budget
also uses the *effective* quantization step from @tbl-s1-enob rather than the
nominal LSB, because the converter's own noise, not its code width, is what
sets the resolution here.
:::

## What to report

1. @tbl-s1-cal and @fig-s1-cal: the fit, with offset and gain and their
   uncertainties.
2. @fig-s1-hist and @tbl-s1-noise: the noise on both channels, in volts and in
   LSB, with the shorted channel identified as the floor.
3. @tbl-s1-enob: effective bits against nominal, and which of the two you
   would quote in a specification, and why.
4. @fig-s1-timing and @tbl-s1-timing: the residual pattern classified as band,
   slope or step, and the number of long intervals.
5. @tbl-s1-hosts and @fig-s1-hosts: the host comparison, with a recommendation
   you can defend from your own numbers.
6. @tbl-s1-budget: the budget, the dominant component, and what you would
   change first to reduce it.
